---
title: "Discordant Genes across various sources of Data"
author: "Sonali Arora, Hamid Bolouri"
date: "December 6, 2018"
output: 
  html_document:
    toc: true
    theme: united
---

## Introduction 

First, we will calculate the Discordant genes and samples for GTEx Data, 
followed by calculation of discordant genes and samples for TCGA data. For a 
gene to be discordant, expression of gene in 1 data set should be more than 
32 TPM (i.e. log2 TPM more than 5) and the log2 fold change should be more 
than 2 (i.e. >4-fold difference in expression). 

## Discordant Genes in GTEx

```{r gtex-calculation}
rm(list=ls())
suppressPackageStartupMessages({
  library(SummarizedExperiment)
  library(Hmisc)
  library(ggplot2)
  library(pheatmap)
  library(RColorBrewer)
  library(eulerr)
  library(UpSetR)
  library(grid)
  library(gridExtra)
})

# folder where S3BUCKET data and github directory are stored. eg: ~/Downloads
bigdir = dirname(getwd())
# github directory eg: ~/Downloads/UncertaintyRNA
git_dir = file.path(bigdir,  "UncertaintyRNA")
# S3 bucket directory eg: ~/Downloads/OriginalTCGAGTExData
s3_dir = file.path(bigdir,  "OriginalTCGAGTExData")

# when you run our RMD files, all results will be stored here. 
# This will essentially remake the "data" subfolder from github repo.
# eg:~/Downloads/data
results_dir = file.path(bigdir, "data")


if(!file.exists( file.path(s3_dir, "SE_objects"))){
  stop("Please go through vignette 3 & 4 to make SE objects or download from S3 bucket")
}
if(!file.exists( file.path( results_dir))){
   system(paste0("mkdir ", results_dir))
}
if(!file.exists( file.path( results_dir,"discordant"))){
  system(paste0("mkdir ", file.path(results_dir, "discordant")))
}

if(!file.exists( file.path( results_dir, "pdf"))){
   system(paste0("mkdir ", file.path(results_dir, "pdf")))
}

if(!file.exists( file.path( results_dir, "tables"))){
   system(paste0("mkdir ", file.path(results_dir, "tables")))
}

tcga_gdc <- get(load( file.path( s3_dir, "SE_objects","tcga_gdc_log2_TPM.RData")))
tcga_mskcc_norm <- get(load( file.path( s3_dir, "SE_objects", "tcga_mskcc_norm_log2_TPM.RData")))
tcga_mskcc_batch <- get(load( file.path( s3_dir, "SE_objects", "tcga_mskcc_batch_log2_TPM.RData")))
tcga_recount2 <- get(load( file.path( s3_dir, "SE_objects", "tcga_recount2_log2_TPM.RData")))
tcga_xena <- get(load( file.path( s3_dir, "SE_objects", "tcga_xena_log2_TPM.RData")))
tcga_piccolo <- get(load( file.path( s3_dir, "SE_objects","tcga_piccolo_log2_TPM.RData")))

gtex_v6 <- get(load( file.path( s3_dir, "SE_objects","gtex_v6_log2_TPM.RData")))
gtex_mskcc_norm <- get(load( file.path( s3_dir, "SE_objects","gtex_mskcc_norm_log2_TPM.RData")))
gtex_mskcc_batch <- get(load( file.path( s3_dir, "SE_objects","gtex_mskcc_batch_log2_TPM.RData")))
gtex_recount2 <- get(load( file.path( s3_dir, "SE_objects", "gtex_recount2_log2_TPM.RData")))
gtex_xena <- get(load( file.path( s3_dir, "SE_objects","gtex_xena_log2_TPM.RData")))


multi_mapping_genes = read.delim(
  file.path(git_dir,"data","extdata","RNAseq_countingErrors_958BadGenes.txt"), 
                                 header=TRUE,  stringsAsFactors = FALSE)
disease = read.delim(
  file.path(git_dir, "data","extdata", "curated_gene_disease_associations.tsv"),
                header=TRUE, stringsAsFactors=FALSE)

gtex_v6_mat = assay(gtex_v6)
mskcc_fpkm_mat=assay(gtex_mskcc_norm)
mskcc_batch_mat=assay(gtex_mskcc_batch)
recount2_mat=assay(gtex_recount2)
xena_mat= assay(gtex_xena)

geneName = rownames(gtex_v6)

m = 5 # pairwise max of 2 vectors.
lf = 2; fold_no = "4fold" #log2(4 +0.01) ; 
#lf = 1; fold_no ="2fold"
#lf=1.584; fold_no = "3fold"

mismatch_genes_GTEX = sapply( 1:  nrow(gtex_v6_mat), function(idx){
  temp_gtex=gtex_v6_mat[idx, ]
  temp_mskcc_fpkm=mskcc_fpkm_mat[idx, ]
  temp_recount2=recount2_mat[idx, ]
  temp_xena=xena_mat[idx, ]
  
  diff_no = length(unique( c(
    which(abs(temp_gtex- temp_xena) >= lf & pmax(temp_gtex, temp_xena) >= m   ),
    which(abs(temp_gtex- temp_recount2) >= lf & pmax(temp_gtex, temp_recount2) >= m ),
    which(abs(temp_gtex- temp_mskcc_fpkm) >= lf & pmax(temp_gtex, temp_mskcc_fpkm) >= m ),
    
    which(abs(temp_xena- temp_recount2) >=  lf & pmax(temp_recount2, temp_xena) >= m ),
    which(abs(temp_xena- temp_mskcc_fpkm) >= lf & pmax(temp_mskcc_fpkm, temp_xena) >= m ),
    
    which(abs(temp_recount2- temp_mskcc_fpkm) >= lf  & pmax(temp_recount2, temp_mskcc_fpkm) >= m)
  )))
  diff_no
})
names(mismatch_genes_GTEX) = geneName
gtex_bad_genes = names(which(mismatch_genes_GTEX >=19))
gtex_allgenes = geneName

mismatch_table_GTEX = sapply(1:nrow(gtex_v6_mat), function(idx){
  temp_gtexV6 = gtex_v6_mat[idx, ]
  temp_mskcc_fpkm = mskcc_fpkm_mat[idx, ]
  temp_recount2 = recount2_mat[idx, ]
  temp_xena = xena_mat[idx, ]

  diffTbl = c(
    length(which(abs(temp_gtexV6- temp_xena) > lf & pmax(temp_gtexV6, temp_xena) > m)),
    length(which(abs(temp_gtexV6- temp_recount2) > lf & pmax(temp_gtexV6, temp_recount2) > m)),
    length(which(abs(temp_gtexV6- temp_mskcc_fpkm) > lf & pmax(temp_gtexV6, temp_mskcc_fpkm) > m)),

    length(which(abs(temp_xena- temp_recount2) > lf & pmax(temp_recount2, temp_xena) > m)),
     length(which(abs(temp_xena- temp_mskcc_fpkm) > lf & pmax(temp_mskcc_fpkm, temp_xena) > m)),

    length(which(abs(temp_recount2- temp_mskcc_fpkm) > lf  & pmax(temp_recount2, temp_mskcc_fpkm) > m))   )
})
mismatch_table_GTEX <- t(mismatch_table_GTEX)
rownames(mismatch_table_GTEX) = geneName
colnames(mismatch_table_GTEX) = c('gtexV6-xena', 'gtexV6-recount', 'gtexV6-mskcc',
                          'xena-recount', 'xena-mskcc', 'recount-mskcc')
gtexTbl <-mismatch_table_GTEX[gtex_bad_genes, ]


mismatchGenesByDataset = sapply( 1:  nrow(gtex_v6_mat), function(idx){
  temp_gtex=gtex_v6_mat[idx, ]
  temp_mskcc_fpkm=mskcc_fpkm_mat[idx, ]
  temp_recount2=recount2_mat[idx, ]
  temp_xena=xena_mat[idx, ]

  gtex_no = length(unique( c(
    which(abs(temp_gtex- temp_xena) > lf & pmax(temp_gtex, temp_xena) > m   ), 
    which(abs(temp_gtex- temp_recount2) > lf & pmax(temp_gtex, temp_recount2) > m ), 
    which(abs(temp_gtex- temp_mskcc_fpkm) > lf & pmax(temp_gtex, temp_mskcc_fpkm) > m )

  )))

  xena_no = length(unique( c(
    which(abs(temp_xena- temp_gtex) > lf  & pmax(temp_gtex, temp_xena) > m ), 
    which(abs(temp_xena- temp_recount2) > lf & pmax(temp_recount2, temp_xena) > m ),
    which(abs(temp_xena- temp_mskcc_fpkm) > lf & pmax(temp_mskcc_fpkm, temp_xena) > m )
  )))


  recount2_no = length(unique( c(

    which(abs(temp_recount2- temp_gtex) > lf   & pmax(temp_recount2, temp_gtex) > m),
    which(abs(temp_recount2- temp_xena) > lf  & pmax(temp_recount2, temp_xena) > m),
    which(abs(temp_recount2- temp_mskcc_fpkm) > lf  & pmax(temp_recount2, temp_mskcc_fpkm) > m)
  )))


  norm_no = length(unique( c(
    which(abs(temp_mskcc_fpkm- temp_gtex) > lf & pmax(temp_mskcc_fpkm, temp_gtex) > m ),
    which(abs(temp_mskcc_fpkm- temp_xena) > lf & pmax(temp_mskcc_fpkm, temp_xena) > m),
    which(abs(temp_mskcc_fpkm- temp_recount2) > lf & pmax(temp_mskcc_fpkm, temp_recount2) > m)
    )))


  c( gtex_no, xena_no, recount2_no,  norm_no) #, batch_no)

})

mismatchGenesByDataset= t(mismatchGenesByDataset)
rownames(mismatchGenesByDataset) = geneName
colnames(mismatchGenesByDataset) = c( "gtex_no", "xena_no", "recount2_no", "norm_no")


gtex_lst =  list(
  gtex_genes = rownames(mismatchGenesByDataset[which( mismatchGenesByDataset[,"gtex_no"] >= 19), ]),
  norm_genes = rownames(mismatchGenesByDataset[which( mismatchGenesByDataset[,"norm_no"] >= 19), ]),
  recount2_genes = rownames(mismatchGenesByDataset[which( mismatchGenesByDataset[,"recount2_no"] >= 19), ]),
  xena_genes = rownames(mismatchGenesByDataset[which( mismatchGenesByDataset[,"xena_no"] >= 19), ])
)

# clear some objects that we don't need: 
rm(gtex_v6)
rm(gtex_mskcc_norm)
rm(gtex_mskcc_batch)
rm(gtex_recount2)
rm(gtex_xena)

rm(gtex_v6_mat)
rm(mskcc_fpkm_mat) 
rm(mskcc_batch_mat) 
rm(recount2_mat)
rm(xena_mat) 

rm(mismatch_genes_GTEX)
rm(mismatch_table_GTEX)

```

## Discordant Genes in TCGA

```{r tcga-calculation}
geneName = rownames(tcga_gdc)

gdc_mat = assay(tcga_gdc)
mskcc_fpkm_mat=assay(tcga_mskcc_norm)
mskcc_batch_mat=assay(tcga_mskcc_batch)
piccolo_mat=assay(tcga_piccolo)
recount2_mat=assay(tcga_recount2)
xena_mat= assay(tcga_xena)

mismatch_genes_TCGA = sapply( 1:  nrow(gdc_mat), function(idx){
  temp_gdc=gdc_mat[idx, ]
  temp_piccolo=piccolo_mat[idx, ]
  temp_mskcc_batch=mskcc_batch_mat[idx, ]
  temp_mskcc_fpkm=mskcc_fpkm_mat[idx, ]
  temp_recount2=recount2_mat[idx, ]
  temp_xena=xena_mat[idx, ]
  
  diff_no = length(unique( c(
    which(abs(temp_gdc- temp_xena) > lf & pmax(temp_gdc, temp_xena) > m   ),
    which(abs(temp_gdc- temp_recount2) > lf & pmax(temp_gdc, temp_recount2) > m ),
    which(abs(temp_gdc- temp_mskcc_fpkm) > lf & pmax(temp_gdc, temp_mskcc_fpkm) > m ),
    which(abs(temp_gdc- temp_piccolo) > lf & pmax(temp_gdc, temp_piccolo) > m ),
    
    which(abs(temp_xena- temp_recount2) > lf & pmax(temp_recount2, temp_xena) > m ),
    which(abs(temp_xena- temp_piccolo) > lf & pmax(temp_piccolo, temp_xena) > m ),
    which(abs(temp_xena- temp_mskcc_fpkm) > lf & pmax(temp_mskcc_fpkm, temp_xena) > m ),
    
    which(abs(temp_recount2- temp_piccolo) > lf  & pmax(temp_recount2, temp_piccolo) > m),
    which(abs(temp_recount2- temp_mskcc_fpkm) > lf  & pmax(temp_recount2, temp_mskcc_fpkm) > m),
    
   which(abs(temp_mskcc_fpkm- temp_piccolo) > lf & pmax(temp_mskcc_fpkm, temp_piccolo) > m)
    
  )))
  diff_no
})
names(mismatch_genes_TCGA) = geneName
tcga_bad_genes= names(which(mismatch_genes_TCGA > 48))
tcga_allgenes = geneName

mismatch_table_TCGA = sapply(1:nrow(gdc_mat), function(idx){
  temp_gdc = gdc_mat[idx, ]
  temp_piccolo = piccolo_mat[idx, ]
  temp_mskcc_batch = mskcc_batch_mat[idx, ]
  temp_mskcc_fpkm = mskcc_fpkm_mat[idx, ]
  temp_recount2 = recount2_mat[idx, ]
  temp_xena = xena_mat[idx, ]

  diffTbl = c(
    length(which(abs(temp_gdc- temp_xena) > lf & pmax(temp_gdc, temp_xena) > m)),
    length(which(abs(temp_gdc- temp_recount2) > lf & pmax(temp_gdc, temp_recount2) > m)),
    length(which(abs(temp_gdc- temp_mskcc_fpkm) > lf & pmax(temp_gdc, temp_mskcc_fpkm) > m)),
    length(which(abs(temp_gdc- temp_piccolo) > lf & pmax(temp_gdc, temp_piccolo) > m)),

    length(which(abs(temp_xena- temp_recount2) > lf & pmax(temp_recount2, temp_xena) > m)),
    length(which(abs(temp_xena- temp_piccolo) > lf & pmax(temp_piccolo, temp_xena) > m)),
    length(which(abs(temp_xena- temp_mskcc_fpkm) > lf & pmax(temp_mskcc_fpkm, temp_xena) > m)),

    length(which(abs(temp_recount2- temp_piccolo) > lf  & pmax(temp_recount2, temp_piccolo) > m)),
    length(which(abs(temp_recount2- temp_mskcc_fpkm) > lf  & pmax(temp_recount2, temp_mskcc_fpkm) > m)),

    length(which(abs(temp_mskcc_fpkm- temp_piccolo) > lf & pmax(temp_mskcc_fpkm, temp_piccolo) > m))
  )
 
})
mismatch_table_TCGA <- t(mismatch_table_TCGA)
rownames(mismatch_table_TCGA) = geneName
colnames(mismatch_table_TCGA) = c('gdc-xena', 'gdc-recount', 'gdc-mskcc', 'gdc-piccolo',
                                'xena-recount', 'xena-piccolo', 'xena-mskcc',
                                'recount-piccolo', 'recount-mskcc', 'mskcc-piccolo')
tcgaTbl <-mismatch_table_TCGA[tcga_bad_genes, ]



mismatchGenesByDataset = sapply( 1:  nrow(gdc_mat), function(idx){
  temp_gdc=gdc_mat[idx, ]
  temp_piccolo=piccolo_mat[idx, ]
  temp_mskcc_fpkm=mskcc_fpkm_mat[idx, ]
  temp_recount2=recount2_mat[idx, ]
  temp_xena=xena_mat[idx, ]

  gdc_no = length(unique( c(
    which(abs(temp_gdc- temp_xena) > lf & pmax(temp_gdc, temp_xena) > m   ), 
    which(abs(temp_gdc- temp_recount2) > lf & pmax(temp_gdc, temp_recount2) > m ),     
    which(abs(temp_gdc- temp_mskcc_fpkm) > lf & pmax(temp_gdc, temp_mskcc_fpkm) > m ),
    which(abs(temp_gdc- temp_piccolo) > lf & pmax(temp_gdc, temp_piccolo) > m )
    
  )))

  xena_no = length(unique( c(
    which(abs(temp_xena- temp_gdc) > lf  & pmax(temp_gdc, temp_xena) > m ), 
    which(abs(temp_xena- temp_recount2) > lf & pmax(temp_recount2, temp_xena) > m ), 
    which(abs(temp_xena- temp_piccolo) > lf & pmax(temp_piccolo, temp_xena) > m ),
    which(abs(temp_xena- temp_mskcc_fpkm) > lf & pmax(temp_mskcc_fpkm, temp_xena) > m )
  )))


  recount2_no = length(unique( c(

    which(abs(temp_recount2- temp_gdc) > lf   & pmax(temp_recount2, temp_gdc) > m),
    which(abs(temp_recount2- temp_xena) > lf  & pmax(temp_recount2, temp_xena) > m),
    which(abs(temp_recount2- temp_piccolo) > lf  & pmax(temp_recount2, temp_piccolo) > m),
    which(abs(temp_recount2- temp_mskcc_fpkm) > lf  & pmax(temp_recount2, temp_mskcc_fpkm) > m)
  )))

  norm_no = length(unique( c(
    which(abs(temp_mskcc_fpkm- temp_gdc) > lf & pmax(temp_mskcc_fpkm, temp_gdc) > m ),
    which(abs(temp_mskcc_fpkm- temp_xena) > lf & pmax(temp_mskcc_fpkm, temp_xena) > m),
    which(abs(temp_mskcc_fpkm- temp_piccolo) > lf & pmax(temp_mskcc_fpkm, temp_piccolo) > m),
    which(abs(temp_mskcc_fpkm- temp_recount2) > lf & pmax(temp_mskcc_fpkm, temp_recount2) > m)
  )))

  piccolo_no = length(unique( c(

    which(abs(temp_piccolo- temp_gdc) > lf & pmax(temp_piccolo, temp_gdc) > m),
    which(abs(temp_piccolo- temp_xena) > lf & pmax(temp_piccolo, temp_xena) > m),
    which(abs(temp_piccolo- temp_mskcc_fpkm) > lf & pmax(temp_piccolo, temp_mskcc_fpkm) > m),
    which(abs(temp_piccolo- temp_recount2) > lf & pmax(temp_piccolo, temp_recount2) > m)
  )))
  c( gdc_no, xena_no, recount2_no, piccolo_no, norm_no) 

})

mismatchGenesByDataset= t(mismatchGenesByDataset)
rownames(mismatchGenesByDataset) = geneName
colnames(mismatchGenesByDataset) = c( "gdc_no", "xena_no", 
                                      "recount2_no", "piccolo_no", "norm_no") 


tcga_lst =  list(
  gdc_genes = rownames(mismatchGenesByDataset[which( mismatchGenesByDataset[,"gdc_no"] > 48), ]),
  norm_genes = rownames(mismatchGenesByDataset[which( mismatchGenesByDataset[,"norm_no"] > 48), ]),
  recount2_genes = rownames(mismatchGenesByDataset[which( mismatchGenesByDataset[,"recount2_no"] > 48), ]),
  xena_genes = rownames(mismatchGenesByDataset[which( mismatchGenesByDataset[,"xena_no"] > 48), ]),
  piccolo_genes = rownames(mismatchGenesByDataset[which( mismatchGenesByDataset[,"piccolo_no"] > 48), ])
)

# clear some objects and free memory 
rm(gdc_mat)
rm(piccolo_mat)
rm(mskcc_batch_mat)
rm(mskcc_fpkm_mat)
rm(recount2_mat)
rm(xena_mat)

rm(tcga_gdc)
rm(tcga_mskcc_norm) 
rm(tcga_mskcc_batch) 
rm(tcga_recount2)
rm(tcga_xena)
rm(tcga_piccolo)

rm(mismatch_genes_TCGA)
rm(mismatch_table_TCGA)

```

## Calculate Percentage of Discordant Genes

```{r}
length(gtex_bad_genes)
length(gtex_allgenes)
(length(gtex_bad_genes)/ length(gtex_allgenes)) *100

length(tcga_bad_genes)
length(tcga_allgenes)
(length(tcga_bad_genes)/ 16109)*100

commongenes = unique( c( tcga_allgenes, gtex_allgenes))
length(commongenes)

bad_genes  = unique( c( tcga_bad_genes, gtex_bad_genes))
length(bad_genes)
(length(bad_genes)/16730 )* 100

multimapped_discordant_genes = intersect(bad_genes, multi_mapping_genes[,3])
length(multimapped_discordant_genes)
(length(multimapped_discordant_genes)/ length(bad_genes))*100

```

## Disease causing genes that are discordant

```{r}
sp = split(disease, disease[,2])
dres = lapply(sp, function(x){
  dname = paste(x[, "diseaseName"], collapse = ", ")
  dcode = paste(x[, "diseaseId"], collapse = ", ")
  x= x[1, ]
  x$diseaseName = dname
  x$diseaseId = dcode
  x
})
dres = do.call(rbind, dres)
disease_genes = dres[, c("geneId", "geneSymbol", "diseaseId", "diseaseName")]

length(intersect(disease_genes[,2], bad_genes)) 

cancer_genes = disease_genes[ match(intersect(disease_genes[,2], bad_genes), disease_genes[,2]), ]
cancer_genes = cancer_genes[order(cancer_genes[,2], decreasing=FALSE), ]
cancertype = disease[match( as.character( cancer_genes[,2]), disease[,2]), ]
cancertype = cancertype[, c(2, 4)]
```

## Save Results 

```{r}
save(gtexTbl, 
  file =  file.path( results_dir, "discordant", 
                     paste0("Heatmap_gtexTbl",fold_no,".RData")))
save(tcgaTbl, 
  file = file.path( results_dir, "discordant", 
                    paste0("Heatmap_tcgaTbl",fold_no,".RData")))

save(tcga_lst,
  file= file.path( results_dir, "discordant",
                   paste0("TCGA_genes_diff_individual_sources",fold_no,".RData")))
save(gtex_lst, 
  file=file.path( results_dir, "discordant", 
                  paste("GTEX_genes_diff_individual_sources",fold_no,".RData")))

write.table( bad_genes, 
  file.path( results_dir, "discordant", paste0("overall_bad_genes_",fold_no, ".txt")), 
  sep="\t", quote=FALSE, row.names=FALSE, col.names=FALSE)

write.table( tcga_bad_genes, 
  file.path( results_dir, "discordant", paste0("tcga_bad_genes_",fold_no, ".txt")), 
  sep="\t", quote=FALSE, row.names=FALSE, col.names=FALSE)

write.table( gtex_bad_genes, 
  file.path(results_dir, "discordant", paste0("gtex_bad_genes_",fold_no, ".txt")), 
  sep="\t", quote=FALSE, row.names=FALSE, col.names=FALSE)

write.table( multimapped_discordant_genes, 
  file.path(results_dir, "discordant", paste0("multimapped_discordant_genes_",fold_no, ".txt")),
  sep="\t", quote=FALSE, row.names=FALSE, col.names=FALSE)

write.table(cancertype, 
  file.path(results_dir, "discordant", paste0("disease_discordant_genes_",fold_no,".txt")), 
  sep="\t", quote = FALSE, row.names=FALSE, col.names=TRUE)

write.table(gtexTbl,
  file.path( results_dir, "tables", "Supp_Table_GTEX_discordant_samples.txt"), 
  sep="\t", quote=FALSE, row.names=TRUE, col.names=TRUE)

write.table(tcgaTbl, 
  file.path( results_dir, "tables", "Supp_Table_TCGA_discordant_samples.txt"), 
  sep="\t", quote=FALSE, row.names=TRUE, col.names=TRUE)
```

