---
title: "Fold change using abundances from different pipelines"
author: "Sonali Arora, Hamid Bolouri"
date: "December 7, 2018"
output: 
  html_document:
    toc: true
    theme: united
---

## Introduction

In this vignette, we want to explore if expression differences between pairs of samples 
for any given gene vary depending on data source. More concretely, 
if we compared the fold change in a gene of interest between two samples, 
across various pipelines, would we see a difference. 

## Calculate Fold Change Differences in TCGA Data 

```{r eval=FALSE}
rm(list=ls())
suppressPackageStartupMessages({
  library(SummarizedExperiment)
  library(Hmisc)
  library(ggplot2)
  library(pheatmap)
  library(RColorBrewer)
  library(eulerr)
  library(UpSetR)
  library(grid)
  library(gridExtra)
})

# folder where S3BUCKET data and github directory are stored. eg: ~/Downloads
bigdir = dirname(getwd())
# github directory eg: ~/Downloads/UncertaintyRNA
git_dir = file.path(bigdir,  "UncertaintyRNA")
# S3 bucket directory eg: ~/Downloads/OriginalTCGAGTExData
s3_dir = file.path(bigdir,  "OriginalTCGAGTExData")

# when you run our RMD files, all results will be stored here. 
# This will essentially remake the "data" subfolder from github repo.
# eg:~/Downloads/data
results_dir = file.path(bigdir, "data")


if(!file.exists( file.path(s3_dir, "SE_objects"))){
  stop("Please go through vignette 3 & 4 to make SE objects or download from S3 bucket")
}
if(!file.exists( file.path( results_dir))){
  system(paste0("mkdir ", results_dir))
}
if(!file.exists( file.path( results_dir,"discordant"))){
  system(paste0("mkdir ", file.path(results_dir, "discordant")))
}

if(!file.exists( file.path( results_dir, "pdf"))){
  system(paste0("mkdir ", file.path(results_dir, "pdf")))
}

if(!file.exists( file.path( results_dir, "tables"))){
  system(paste0("mkdir ", file.path(results_dir, "tables")))
}

tcga_gdc <- get(load( file.path( s3_dir, "SE_objects","tcga_gdc_log2_TPM.RData")))
tcga_mskcc_norm <- get(load( file.path( s3_dir, "SE_objects", "tcga_mskcc_norm_log2_TPM.RData")))
tcga_recount2 <- get(load( file.path( s3_dir, "SE_objects", "tcga_recount2_log2_TPM.RData")))
tcga_xena <- get(load( file.path( s3_dir, "SE_objects", "tcga_xena_log2_TPM.RData")))
tcga_piccolo <- get(load( file.path( s3_dir, "SE_objects","tcga_piccolo_log2_TPM.RData")))

geneName = rownames(tcga_gdc)

m = 5 # pairwise max of 2 vectors.
lf = 2; fold_no = "4fold"

# read in previously calculated discordant genes and subset all se objects
# to contain only discordant genes.
discordant_genes = read.delim(
  file.path(git_dir, "data", "discordant", "tcga_bad_genes_4fold.txt"),
  header=FALSE, stringsAsFactors = FALSE)[,1]
length(discordant_genes)

geneName = rownames(tcga_gdc)
dis_idx = match(discordant_genes, geneName)

tcga_gdc = tcga_gdc[dis_idx, ]
tcga_mskcc_norm = tcga_mskcc_norm[dis_idx, ]
tcga_piccolo = tcga_piccolo[dis_idx, ]
tcga_recount2 = tcga_recount2[dis_idx, ]
tcga_xena = tcga_xena[dis_idx, ]

# extract data from each se object
gdc_mat = assay(tcga_gdc)
mskcc_norm_mat=assay(tcga_mskcc_norm)
piccolo_mat=assay(tcga_piccolo)
recount2_mat=assay(tcga_recount2)
xena_mat= assay(tcga_xena)

# quick calculation : find union of discordant samples per gene
diffsamples_TCGA = sapply(1:nrow(gdc_mat), function(idx){
  temp_gdc = gdc_mat[idx, ]
  temp_piccolo = piccolo_mat[idx, ]
  temp_mskcc_norm= mskcc_norm_mat[idx, ]
  temp_recount2 = recount2_mat[idx, ]
  temp_xena = xena_mat[idx, ]
  
  diffsamples = unique(c(
    (which(abs(temp_gdc- temp_xena) > lf & pmax(temp_gdc, temp_xena) > m)),
    (which(abs(temp_gdc- temp_recount2) > lf & pmax(temp_gdc, temp_recount2) > m)),
    (which(abs(temp_gdc- temp_mskcc_norm) > lf & pmax(temp_gdc, temp_mskcc_norm) > m)),
    (which(abs(temp_gdc- temp_piccolo) > lf & pmax(temp_gdc, temp_piccolo) > m)),
    (which(abs(temp_xena- temp_recount2) > lf & pmax(temp_recount2, temp_xena) > m)),
    (which(abs(temp_xena- temp_piccolo) > lf & pmax(temp_piccolo, temp_xena) > m)),
    (which(abs(temp_xena- temp_mskcc_norm) > lf & pmax(temp_mskcc_norm, temp_xena) > m)),
    (which(abs(temp_recount2- temp_piccolo) > lf  & pmax(temp_recount2, temp_piccolo) > m)),
    (which(abs(temp_recount2- temp_mskcc_norm) > lf  & pmax(temp_recount2, temp_mskcc_norm) > m)),
    (which(abs(temp_mskcc_norm- temp_piccolo) > lf & pmax(temp_mskcc_norm, temp_piccolo) > m))
  ))
})

# note : 
# for some genes, all 4800 samples are discordant across pipelines.
# save these to make pairwise scatter plots ( supp figures)
super_bad <- which(sapply(diffsamples_TCGA, length)==4800)
badGenes <-  discordant_genes[super_bad]
write.table(badGenes, file.path(results_dir, "discordant",
            "TCGA_bad_genes_samples_varying_fc_across_pipelines.txt"),
            sep="\t", quote=FALSE, row.names=FALSE, col.names=TRUE)


thresh <- 0
selGenes <- rownames(tcga_gdc)
result <- lapply( selGenes, function(i){
  gene_name = i
  idx = grep(paste0("^",gene_name,"$"), discordant_genes)
  samples =diffsamples_TCGA[[idx]]
  
  message(i, ":", length(samples))
  
  temp_gdc = gdc_mat[idx, samples]
  temp_piccolo = piccolo_mat[idx, samples]
  temp_mskcc_norm = mskcc_norm_mat[idx, samples ]
  temp_recount2 = recount2_mat[idx, samples]
  temp_xena = xena_mat[idx, samples]
  
  temp <- cbind( temp_gdc,temp_piccolo , temp_mskcc_norm, temp_recount2, 
                 temp_xena )
  temp[which(temp < thresh)] <- NA
  
  gdc_ap1 <- as.vector(dist(temp[,1], method = "manhattan"))
  pic_ap1 <- as.vector(dist(temp[,2], method = "manhattan"))
  mskcc_ap1 <- as.vector(dist(temp[,3], method = "manhattan"))
  rec_ap1 <- as.vector(dist(temp[,4], method = "manhattan"))
  xena_ap1 <- as.vector(dist(temp[,5], method = "manhattan"))
  
  gene_mat <- cbind(gdc_ap1, pic_ap1, mskcc_ap1, rec_ap1, xena_ap1)
  colnames(gene_mat) = c("GDC", "Piccolo", "MSKCC", "Recount2", "Xena/Toil")
  
  #in some cases all pipelines report expression levels less than thresh
  rm_idx <- which(apply(gene_mat, 1 , function(x) all(is.na(x)) ))
  if(length(rm_idx)!=0){
    gene_mat <- gene_mat[ -rm_idx, ]
  }
  
  gene_max <- apply(gene_mat, 1, function(x) max(x, na.rm = TRUE))
  gene_min <- apply(gene_mat, 1,  function(x) min(x, na.rm = TRUE))
  max_min <- gene_max - gene_min
  
  ans1 <- max(max_min)
  
  ans2 <- min(apply(gene_mat,1,  function(x){
   x = na.omit(x) # removes all NA values.
   if(length(x)==1){
     ans <- x
   }else{
     min(sort(abs(diff(x))))
   }
  }))
  
  c(ans1, ans2)
})

result2 <- do.call( rbind, result)
rownames(result2) <- selGenes
result2 <- result2[order(result2[,1], decreasing = TRUE), ]
colnames(result2) <- c("max", "min")
result2 <- data.frame(cbind(gene= rownames(result2), result2))

write.table(result2, 
            file.path(results_dir, "discordant",  "TCGA_fc_table_all_samples.txt"),
            sep="\t", quote=FALSE, row.names=FALSE, col.names=TRUE)
```

## Calculate Fold Change Differences in GTEx Data

```{r eval=FALSE}
rm(list=ls())
library(SummarizedExperiment)

# folder where S3BUCKET data and github directory are stored. eg: ~/Downloads
bigdir = dirname(getwd())
# github directory eg: ~/Downloads/UncertaintyRNA
git_dir = file.path(bigdir,  "UncertaintyRNA")
# S3 bucket directory eg: ~/Downloads/OriginalTCGAGTExData
s3_dir = file.path(bigdir,  "OriginalTCGAGTExData")

# when you run our RMD files, all results will be stored here. 
# This will essentially remake the "data" subfolder from github repo.
# eg:~/Downloads/data
results_dir = file.path(bigdir, "data")

gtex_v6 <- get(load( file.path( s3_dir, "SE_objects","gtex_v6_log2_TPM.RData")))
gtex_mskcc_norm <- get(load( file.path( s3_dir, "SE_objects","gtex_mskcc_norm_log2_TPM.RData")))
gtex_recount2 <- get(load( file.path( s3_dir, "SE_objects", "gtex_recount2_log2_TPM.RData")))
gtex_xena <- get(load( file.path( s3_dir, "SE_objects","gtex_xena_log2_TPM.RData")))

discordant_genes = read.delim(
      file.path(git_dir, "data", "discordant", "gtex_bad_genes_4fold.txt"),
      header=FALSE, stringsAsFactors = FALSE)[,1]
length(discordant_genes)
geneName = rownames(gtex_v6)
dis_idx = match(discordant_genes, geneName)

gtex_v6 = gtex_v6[dis_idx, ]
gtex_mskcc_norm = gtex_mskcc_norm[dis_idx, ]
gtex_recount2 = gtex_recount2[dis_idx, ]
gtex_xena = gtex_xena[dis_idx, ]

gtex_v6_mat = assay(gtex_v6)
mskcc_norm_mat=assay(gtex_mskcc_norm)
recount2_mat=assay(gtex_recount2)
xena_mat= assay(gtex_xena)

m = 5 # pairwise max of 2 vectors.
lf = 2; fold_no = "4fold"

# quick calculation : find union of discordant samples per gene
diffsamples_GTEX = sapply(1:nrow(gtex_v6), function(idx){
  temp_gtex=gtex_v6_mat[idx, ]
  temp_mskcc_norm=mskcc_norm_mat[idx, ]
  temp_recount2=recount2_mat[idx, ]
  temp_xena=xena_mat[idx, ]
  
  diffsamples = unique(c(
    (which(abs(temp_gtex- temp_xena) > lf & pmax(temp_gtex, temp_xena) > m)),
    (which(abs(temp_gtex- temp_recount2) > lf & pmax(temp_gtex, temp_recount2) > m)),
    (which(abs(temp_gtex- temp_mskcc_norm) > lf & pmax(temp_gtex, temp_mskcc_norm) > m)),
    
    (which(abs(temp_xena- temp_recount2) > lf & pmax(temp_recount2, temp_xena) > m)),
    (which(abs(temp_xena- temp_mskcc_norm) > lf & pmax(temp_mskcc_norm, temp_xena) > m)),
    (which(abs(temp_recount2- temp_mskcc_norm) > lf  & pmax(temp_recount2, temp_mskcc_norm) > m))
  ))
})

# note : 
# for some genes, all 1890 samples are discordant.
# save these to make pairwise scatter plots ( supp figures)
super_bad <- which(sapply(diffsamples_GTEX, length)==1890)
badGenes <-  discordant_genes[super_bad]
write.table(badGenes, file.path(results_dir, "discordant",  
            "GTEX_bad_genes_samples_varying_fc_across_pipelines.txt"),
            sep="\t", quote=FALSE, row.names=FALSE, col.names=TRUE)


thresh <- 0
selGenes <- rownames(gtex_v6_mat)
result <- lapply( selGenes, function(i){
  gene_name = i
  idx = grep(paste0("^",gene_name,"$"), selGenes)
  samples =diffsamples_GTEX[[idx]]
  gene_mat = matrix( nrow = length(samples)-1, ncol=4 )
  
  message(i, ":", length(samples))
  
  temp_gtex=gtex_v6_mat[idx, samples]
  temp_mskcc_norm=mskcc_norm_mat[idx, samples ]
  temp_recount2=recount2_mat[idx, samples]
  temp_xena=xena_mat[idx, samples]
  
  temp <- cbind( temp_gtex, temp_mskcc_norm, temp_recount2, temp_xena)
  temp[which(temp < thresh)] <- NA
  
  gtex_ap1 <- as.vector(dist(temp[,1], method = "manhattan"))
  mskcc_ap1 <- as.vector(dist(temp[,2], method = "manhattan"))
  rec_ap1 <- as.vector(dist(temp[,3], method = "manhattan"))
  xena_ap1 <- as.vector(dist(temp[,4], method = "manhattan"))
  
  gene_mat <- cbind(gtex_ap1, mskcc_ap1, rec_ap1, xena_ap1)
  colnames(gene_mat) = c("GTEX", "MSKCC", "Recount2", "Xena/Toil")
  
  # in some cases all pipelines have expression values less than thresh
  rm_idx <- which(apply(gene_mat, 1 , function(x) all(is.na(x)) ))
  if(length(rm_idx)!=0){
    gene_mat <- gene_mat[ -rm_idx, ]
  }
  
  gene_max <- apply(gene_mat, 1, function(x) max(x, na.rm = TRUE))
  gene_min <- apply(gene_mat, 1,  function(x) min(x, na.rm = TRUE))
  max_min <- gene_max - gene_min
  
  ans1 <- max(max_min)
  
  # min apprach 
  ans2 <- min(apply(gene_mat,1,  function(x){
   x = na.omit(x) # removes all NA values.
   if(length(x)==1){
     ans <- x
   }else{
     min(abs(diff(sort(x))))
   }
  }))
  c(ans1, ans2) 
})

result2 <- do.call( rbind, result)
rownames(result2) <- selGenes
result2 <- result2[order(result2[,1], decreasing = TRUE), ]
colnames(result2) <- c("max", "min")
result2 <- data.frame(cbind(gene= rownames(result2), result2))

write.table(result2, 
            file.path(results_dir, "discordant", "GTEX_fc_table_all_samples.txt"),
            sep="\t", quote=FALSE, row.names=FALSE, col.names=TRUE)
```
