---
title: "Supplemental Tables"
author: "Sonali Arora, Hamid Bolouri"
date: "October 1, 2018"
output: 
  html_document:
    toc: true
    theme: united
---

## Introduction

In this vignette, we calculate various Supplemental Tables, as seen in our Paper.

## mRNA Correlations - TCGA

```{r}
rm(list=ls())
suppressPackageStartupMessages({
  library(SummarizedExperiment)
  library(Hmisc)
  library(ggplot2)
  library(pheatmap)
  library(RColorBrewer)
})

# folder where S3BUCKET data and github directory are stored. eg: ~/Downloads
bigdir = dirname(getwd())
# github directory eg: ~/Downloads/UncertainityRNA
git_dir = file.path(bigdir,  "UncertainityRNA")
# S3 bucket directory eg: ~/Downloads/OriginalTCGAGTExData
s3_dir = file.path(bigdir,  "OriginalTCGAGTExData")

# when you run our RMD files, all results will be stored here. 
# This will essentially remake the "data" subfolder from github repo.
# eg:~/Downloads/data
results_dir = file.path(bigdir, "data")

if(!file.exists( s3_dir)){
  stop("Please download from S3 bucket")
}

if(!file.exists( file.path(s3_dir, "SE_objects"))){
  stop("Please go through vignette 3 & 4 to make SE objects or download from S3 bucket")
}
if(!file.exists( file.path( results_dir))){
   system(paste0("mkdir ", results_dir))
}
if(!file.exists( file.path( results_dir, "tables"))){
   system(paste0("mkdir ", file.path(results_dir, "tables")))
}

tcga_gdc <- get(load( file.path( s3_dir, "SE_objects","tcga_gdc_log2_TPM.RData")))
tcga_mskcc_norm <- get(load( file.path( s3_dir, "SE_objects", "tcga_mskcc_norm_log2_TPM.RData")))
tcga_mskcc_batch <- get(load( file.path( s3_dir, "SE_objects", "tcga_mskcc_batch_log2_TPM.RData")))
tcga_recount2 <- get(load( file.path( s3_dir, "SE_objects", "tcga_recount2_log2_TPM.RData")))
tcga_xena <- get(load( file.path( s3_dir, "SE_objects", "tcga_xena_log2_TPM.RData")))
tcga_piccolo <- get(load( file.path( s3_dir, "SE_objects","tcga_piccolo_log2_TPM.RData")))

gdc_mat = assay(tcga_gdc)
mskcc_norm_mat=assay(tcga_mskcc_norm)
mskcc_batch_mat=assay(tcga_mskcc_batch)
piccolo_mat=assay(tcga_piccolo)
recount2_mat=assay(tcga_recount2)
xena_mat= assay(tcga_xena)

genes=rownames(gdc_mat)
mat = matrix(nrow=length(genes), ncol = 5)
for(idx in 1:length(genes)){
  geneName = genes[idx]
  temp_gdc=gdc_mat[idx, ]
  temp_xena=xena_mat[idx, ]
  temp_gse62944=piccolo_mat[idx, ]
  temp_recount2=recount2_mat[idx, ]
  temp_mskcc_norm=mskcc_norm_mat[idx, ]
  temp_mskcc_batch=mskcc_batch_mat[idx, ]
  
  r1 <- rcorr(cbind(temp_gdc, temp_xena, temp_gse62944,temp_recount2,
                    temp_mskcc_norm, temp_mskcc_batch ), type="pearson")
  
  corr_xena = r1$r[1,2]
  corr_gse62944= r1$r[1,3]
  corr_recount2= r1$r[1,4]
  corr_mskcc_norm= r1$r[1,5]
  corr_mskcc_batch= r1$r[1,6]
  
  temp = c(corr_xena, corr_gse62944, corr_recount2,
           corr_mskcc_norm, corr_mskcc_batch)
  
  mat[idx, ] = temp
}
rownames(mat)=genes

colnames(mat) = c("corr_XENA_Toil", "corr_Piccolo_Lab", "corr_Recount2",
                "corr_MSKCC_norm", "corr_MSKCC_BATCH")

mat = cbind(gene=genes, mat)
write.table(mat, file.path(results_dir, "tables", 
            "Supp_Table_TCGA_Correlations_all_6_Datasets_log2_TPM.txt"),
            sep="\t", quote=FALSE, row.names=FALSE, col.names=TRUE)
```

## mRNA Correlations - GTEx

```{r}
rm(list=ls())
suppressPackageStartupMessages({
  library(SummarizedExperiment)
  library(Hmisc)
  library(ggplot2)
  library(pheatmap)
  library(RColorBrewer)
})

# folder where S3BUCKET data and github directory are stored. eg: ~/Downloads
bigdir = dirname(getwd())
# github directory eg: ~/Downloads/UncertainityRNA
git_dir = file.path(bigdir,  "UncertainityRNA")
# S3 bucket directory eg: ~/Downloads/OriginalTCGAGTExData
s3_dir = file.path(bigdir,  "OriginalTCGAGTExData")

# when you run our RMD files, all results will be stored here. 
# This will essentially remake the "data" subfolder from github repo.
# eg:~/Downloads/data
results_dir = file.path(bigdir, "data")


if(!file.exists(s3_dir)){
  stop("Please download from S3 bucket")
}
if(!file.exists(file.path(s3_dir, "SE_objects"))){
  stop("Please go through vignette 3 & 4 to make SE objects or download from S3 bucket")
}
if(!file.exists( file.path( results_dir))){
   system(paste0("mkdir ", results_dir))
}
if(!file.exists( file.path( results_dir, "tables"))){
   system(paste0("mkdir ", file.path(results_dir, "tables")))
}


gtex_v6 <- get(load( file.path( s3_dir, "SE_objects","gtex_v6_log2_TPM.RData")))
gtex_mskcc_norm <- get(load( file.path( s3_dir, "SE_objects","gtex_mskcc_norm_log2_TPM.RData")))
gtex_mskcc_batch <- get(load( file.path( s3_dir, "SE_objects","gtex_mskcc_batch_log2_TPM.RData")))
gtex_recount2 <- get(load( file.path( s3_dir, "SE_objects", "gtex_recount2_log2_TPM.RData")))
gtex_xena <- get(load( file.path( s3_dir, "SE_objects","gtex_xena_log2_TPM.RData")))

gtex_v6_mat = assay(gtex_v6)
mskcc_norm_mat=assay(gtex_mskcc_norm)
mskcc_batch_mat=assay(gtex_mskcc_batch)
recount2_mat=assay(gtex_recount2)
xena_mat= assay(gtex_xena)

genes=rownames(gtex_v6_mat)
mat = matrix(nrow=length(genes), ncol = 4)
for(idx in 1:length(genes)){

  temp_gtex_v6=gtex_v6_mat[idx, ]
  temp_xena=xena_mat[idx, ]
  temp_recount2=recount2_mat[idx, ]
  temp_mskcc_norm=mskcc_norm_mat[idx, ]
  temp_mskcc_batch=mskcc_batch_mat[idx, ]

  r1 <- rcorr(cbind(temp_gtex_v6, temp_xena, temp_recount2,
                    temp_mskcc_norm, temp_mskcc_batch ), type="pearson")

  corr_xena = r1$r[1,2]
  corr_recount2= r1$r[1,3]
  corr_mskcc_norm= r1$r[1,4]
  corr_mskcc_batch= r1$r[1,5]

  temp = c(corr_xena, corr_recount2,
           corr_mskcc_norm, corr_mskcc_batch)

  mat[idx, ] = temp
}
rownames(mat)=genes

colnames(mat) = c("corr_XENA_Toil",  "corr_Recount2",
                  "corr_MSKCC_Norm", "corr_MSKCC_BATCH")
mat = cbind(gene=genes, mat)
write.table(mat, file.path(results_dir, "tables", 
            "Supp_Table_GTEX_Correlations_all_6_Datasets_log2_TPM.txt"),
            sep="\t", quote=FALSE, row.names=FALSE, col.names=TRUE)
```


## Protein-mRNA correlations

```{r}
rm(list=ls())

suppressPackageStartupMessages({
  library(SummarizedExperiment)
  library(Hmisc)
  library(ggplot2)
  library(pheatmap)
  library(RColorBrewer)
  library(gridExtra)
  library(grid)
  library(beeswarm)
})

# folder where S3BUCKET data and github directory are stored. eg: ~/Downloads
bigdir = dirname(getwd())
# github directory eg: ~/Downloads/UncertainityRNA
git_dir = file.path(bigdir,  "UncertainityRNA")
# S3 bucket directory eg: ~/Downloads/OriginalTCGAGTExData
s3_dir = file.path(bigdir,  "OriginalTCGAGTExData")

# when you run our RMD files, all results will be stored here. 
# This will essentially remake the "data" subfolder from github repo.
# eg:~/Downloads/data
results_dir = file.path(bigdir, "data")

if(!file.exists( s3_dir)){
  stop("Please download from S3 bucket")
}
if(!file.exists(file.path(s3_dir, "SE_objects"))){
  stop("Please go through vignette 3 & 4 to make SE objects or download from S3 bucket")
}

tcga_gdc <- get(load( file.path( s3_dir, "SE_objects","tcga_gdc_log2_TPM.RData")))
tcga_mskcc_norm <- get(load( file.path( s3_dir, "SE_objects", "tcga_mskcc_norm_log2_TPM.RData")))
tcga_mskcc_batch <- get(load( file.path( s3_dir, "SE_objects", "tcga_mskcc_batch_log2_TPM.RData")))
tcga_recount2 <- get(load( file.path( s3_dir, "SE_objects", "tcga_recount2_log2_TPM.RData")))
tcga_xena <- get(load( file.path( s3_dir, "SE_objects", "tcga_xena_log2_TPM.RData")))
tcga_piccolo <- get(load( file.path( s3_dir, "SE_objects","tcga_piccolo_log2_TPM.RData")))

tcga_gdc_mat = assay(tcga_gdc)
tcga_mskcc_fpkm_mat=assay(tcga_mskcc_norm)
tcga_mskcc_batch_mat=assay(tcga_mskcc_batch)
tcga_piccolo_mat=assay(tcga_piccolo)
tcga_recount2_mat=assay(tcga_recount2)
tcga_xena_mat= assay(tcga_xena)

rownames(tcga_gdc_mat) = rowRanges(tcga_gdc)$external_gene_name

# read in protein data
protein_dir =file.path( git_dir, "data","extdata")
df = read.delim(file.path(protein_dir, "RPPA_RBN"),
                header=T, stringsAsFactors=FALSE, row.names=1 )
map = read.delim(file.path(protein_dir, "md_anderson_antibodies.probeMap"),
                 header=T, stringsAsFactors=FALSE)

# use the same gene-proteins from each source
temp = lapply( rownames(df), function(x) {
  g1 = unique(map[grep(x,  map[,1]), 2])
  intersect(toupper(g1), toupper( rownames(tcga_gdc_mat)) )[1]

})

ix =  which(sapply(temp, is.na))
df = df[ -ix, ]
common_genes = unlist(temp[-ix])

row_idx = match( common_genes, rownames(tcga_gdc_mat))

tcga_gdc_mat = tcga_gdc_mat[row_idx, ]
tcga_mskcc_fpkm_mat= tcga_mskcc_fpkm_mat[row_idx, ]
tcga_mskcc_batch_mat=tcga_mskcc_batch_mat[row_idx, ]
tcga_piccolo_mat=tcga_piccolo_mat[ row_idx, ]
tcga_recount2_mat=tcga_recount2_mat[ row_idx, ]
tcga_xena_mat= tcga_xena_mat[ row_idx, ]

## should use the same TCGA samples from each source.
colnames(df) = gsub("[.]","-", colnames(df))

col_nms = intersect(colnames(df), substr(colnames(tcga_gdc_mat),1,15))

col_idx1 = match(col_nms, colnames(df))
col_idx2 = match(col_nms,  substr(colnames(tcga_gdc_mat),1,15))
df = df[, col_idx1]

tcga_gdc_mat = tcga_gdc_mat[,col_idx2 ]
tcga_mskcc_fpkm_mat= tcga_mskcc_fpkm_mat[,col_idx2 ]
tcga_mskcc_batch_mat=tcga_mskcc_batch_mat[,col_idx2 ]
tcga_piccolo_mat=tcga_piccolo_mat[ ,col_idx2 ]
tcga_recount2_mat=tcga_recount2_mat[ ,col_idx2 ]
tcga_xena_mat= tcga_xena_mat[ ,col_idx2 ]

mat = matrix(nrow=nrow(tcga_gdc_mat), ncol = 6)
for(idx in 1:nrow(tcga_gdc_mat)){
  geneName = rownames(tcga_gdc_mat)[idx]
  protein = unlist(df[idx, ])


  temp_gdc=tcga_gdc_mat[idx, ]
  temp_gse62944=tcga_piccolo_mat[idx, ]
  temp_mskcc_batch=tcga_mskcc_batch_mat[idx, ]
  temp_mskcc_fpkm=tcga_mskcc_fpkm_mat[idx, ]
  temp_recount2=tcga_recount2_mat[idx, ]
  temp_xena=tcga_xena_mat[idx, ]

  r1 <- rcorr(cbind(protein, temp_gdc, temp_xena, temp_gse62944,temp_recount2,
                    temp_mskcc_fpkm,temp_mskcc_batch ), type="spearman")

  corr_gdc = r1$r[1,2]
  corr_xena = r1$r[1,3]
  corr_gse62944= r1$r[1,4]
  corr_recount2= r1$r[1,5]
  corr_mskcc_fpkm= r1$r[1,6]
  corr_mskcc_batch= r1$r[1,7]

  temp = c( corr_gdc, corr_gse62944, corr_mskcc_batch, corr_mskcc_fpkm,
            corr_recount2,corr_xena)

  mat[idx, ] = temp
}
rownames(mat) = rownames(tcga_gdc_mat)
mat = mat[order(as.numeric(mat[,1]), decreasing =TRUE), ]


colnames(mat) = c("corr_protein_GDC", "corr_protein_PiccoloLab",
                  "corr_protein_mskcc_BATCH",
                  "corr_protein_mskcc_Norm", "corr_protein_Recount2",
                  "corr_protein_Xena_Toil" )
mat2 = cbind(gene = rownames(mat), mat)

write.table(mat2, 
            file.path(results_dir, "tables", 
            "Supp_Table_TCGA_Protein_MRNA_Correlations_all_6_Datasets.txt"),
            sep="\t", quote=FALSE, row.names=FALSE, col.names=TRUE)


```

## Table showing region types for each TCGA and GTEx Sample

Note: This table is not a supplemental table in our paper, it is used internally 
while plotting Figures and Supplemental Figures for the paper. 

```{r}
rm(list=ls())
suppressPackageStartupMessages({
  library(SummarizedExperiment)
})


# folder where S3BUCKET data and github directory are stored. eg: ~/Downloads
bigdir = dirname(getwd())
# github directory eg: ~/Downloads/UncertainityRNA
git_dir = file.path(bigdir,  "UncertainityRNA")
# S3 bucket directory eg: ~/Downloads/OriginalTCGAGTExData
s3_dir = file.path(bigdir,  "OriginalTCGAGTExData")

# when you run our RMD files, all results will be stored here. 
# This will essentially remake the "data" subfolder from github repo.
# eg:~/Downloads/data
results_dir = file.path(bigdir, "data")

if(!file.exists( s3_dir)){
  stop("Please download from S3 bucket")
}
if(!file.exists(file.path(s3_dir, "SE_objects"))){
  stop("Please go through vignette 3 & 4 to make SE objects or download from S3 bucket")
}


tcga_gdc <- get(load( file.path( s3_dir, "SE_objects","tcga_gdc_log2_TPM.RData")))
tcga_types <- colData(tcga_gdc)

gtex_v6 <- get(load( file.path( s3_dir, "SE_objects","gtex_v6_log2_TPM.RData")))
gtex_types <- colData(gtex_v6)

tcga_types <- as.data.frame(tcga_types)

gtex_types <- as.data.frame(gtex_types)
gtex_types <- gtex_types[, c("SAMPID","SMTS", "SMTSD", "SMNABTCH","SMNABTCHT",
  "SMNABTCHD","SMGEBTCH","SMGEBTCHD","SMGEBTCHT")]
colnames(gtex_types)[1:3]= c("sampleName", "sampleGroup","region")

write.table(tcga_types, 
            file.path(results_dir, "tables","Supp_Table_TCGA_Cancer_Types.txt"), 
            sep="\t", quote=FALSE, row.names=FALSE, col.names=TRUE)

write.table(gtex_types, 
            file.path(results_dir, "tables", "Supp_Table_GTEX_Types.txt"), 
            sep="\t", quote=FALSE, row.names=FALSE, col.names=TRUE)
```
