---
title: "Supplemental Figures"
author: "Sonali Arora, Hamid Bolouri"
date: "December 6, 2018"
output: 
  html_document:
    toc: true
    theme: united
---

## Introduction

In this vignette, we calculate various Supplemental Figures, as seen in our Paper.

## Pairwise PCA Plots for individual TCGA and GTEx data 

```{r}
rm(list=ls())

library(grid)
library(gridExtra)
library(ggplot2)

# folder where S3BUCKET data and github directory are stored. eg: ~/Downloads
bigdir = dirname(getwd())
# github directory eg: ~/Downloads/UncertaintyRNA
git_dir = file.path(bigdir,  "UncertaintyRNA")
# S3 bucket directory eg: ~/Downloads/OriginalTCGAGTExData
s3_dir = file.path(bigdir,  "OriginalTCGAGTExData")

# when you run our RMD files, all results will be stored here. 
# This will essentially remake the "data" subfolder from github repo.
# eg:~/Downloads/data
results_dir = file.path(bigdir, "data")

if(!file.exists( file.path( git_dir, "data","pca_data"))){
  stop("Please go through vignette 5 & 6 to calculate PCA or clone from github")
}
if(!file.exists( file.path( results_dir))){
   system(paste0("mkdir ", results_dir))
}
if(!file.exists( file.path( results_dir, "pdf"))){
   system(paste0("mkdir ", file.path(results_dir, "pdf")))
}

tcga_col_lst = c( 
  "Piccolo"= "#E76BF3",
  "MSKCC" = "darkblue",
  "MSKCC Batch" ="#00B0F6" ,
  Recount2 ="grey45" ,
  "Xena/Toil" = "orange",
  "GDC" ="#A3A500")

gtex_col_lst = c( 
  "MSKCC" = "darkblue",
  "MSKCC Batch" ="#00B0F6" ,
  Recount2 ="grey45" ,
  "Xena/Toil" = "orange",
  "GTEx" ="darkgreen")


s1 = 3 # size for points in PCA plot
legend_pt_size =4
plot_title_size = 25
axis_text_size = 25
axis_title_size=25
legend_text_size=20
spacing=0.3
chosen_margin = c(0.5,1,0.5,1)# margins:top,right,bottom,left


tcga_fls1 = c( 
  file.path(git_dir, "data","pca_data", "Supp_Fig1_PCA_Data_GDC_vs_XENA_Toil_TPM.txt"),
  file.path(git_dir,"data","pca_data", "Supp_Fig1_PCA_Data_GDC_vs_Piccolo_Lab_TPM.txt")  ,
  file.path(git_dir,"data","pca_data", "Supp_Fig1_PCA_Data_GDC_vs_Recount2_TPM.txt"),
  file.path(git_dir,"data","pca_data", "Supp_Fig1_PCA_Data_GDC_vs_MSKCC_Norm_TPM.txt") ,
  file.path(git_dir,"data","pca_data", "Supp_Fig1_PCA_Data_GDC_vs_MSKCC_BATCH_TPM.txt")  )

gtex_fls1 = c( file.path(git_dir,"data","pca_data", "Supp_Fig1_PCA_Data_Gtex_v6_vs_XENA_Toil_TPM.txt"),
 file.path(git_dir,"data","pca_data", "Supp_Fig1_PCA_Data_Gtex_v6_vs_Recount2_TPM.txt") ,
 file.path(git_dir,"data","pca_data", "Supp_Fig1_PCA_Data_Gtex_v6_vs_MSKCC_Norm_TPM.txt"),
 file.path(git_dir,"data","pca_data", "Supp_Fig1_PCA_Data_Gtex_v6_vs_MSKCC_BATCH_TPM.txt" ))

percentVar_fls = c( 
  file.path(git_dir,"data","pca_data", "percentVar_GDC_vs_XENA_Toil_TPM.txt"),
  file.path(git_dir,"data","pca_data", "percentVar_GDC_vs_Piccolo_Lab_TPM.txt")  ,
  file.path(git_dir,"data","pca_data", "percentVar_GDC_vs_Recount2_TPM.txt"),
  file.path(git_dir,"data","pca_data", "percentVar_GDC_vs_MSKCC_Norm_TPM.txt"),
  file.path(git_dir,"data","pca_data", "percentVar_GDC_vs_MSKCC_BATCH_TPM.txt"))

gtex_percentVar_list = c(
  file.path(git_dir,"data","pca_data", "percentVar_Gtex_vs_XENA_Toil_TPM.txt"), 
  file.path(git_dir,"data","pca_data", "percentVar_Gtex_vs_Recount2_TPM.txt"), 
  file.path(git_dir,"data","pca_data", "percentVar_Gtex_vs_MSKCC_Norm_TPM.txt"), 
  file.path(git_dir,"data","pca_data", "percentVar_Gtex_vs_MSKCC_BATCH_TPM.txt"))

tcga_res = mapply(function(x,y){
  pc_data1 = read.delim(x, header=T, stringsAsFactors=FALSE)
  pc_data1$Project[which(pc_data1$Project=="Piccolo_Lab")]="Piccolo"
  pc_data1$Project[which(pc_data1$Project=="XENA_Toil")]="Xena/Toil"
  pc_data1$Project[which(pc_data1$Project=="MSKCC_Norm")]="MSKCC"
  pc_data1$Project[which(pc_data1$Project=="MSKCC_BATCH")]="MSKCC Batch"
  pc_data1$Project[which(pc_data1$Project=="TCGA-GDCData")]="GDC"

  other_title = setdiff(unique(pc_data1$Project), "GDC")
  other_color = as.character(tcga_col_lst[other_title])
  gdc_color = as.character(tcga_col_lst["GDC"])
  pc_data1$Project = factor(pc_data1$Project,
                            levels= c("GDC", other_title))
  
  percentVar = read.delim( y, header=FALSE, stringsAsFactors = FALSE)[,1]
  
  p1 = ggplot(pc_data1, aes(PC1, PC2, color=Project)) +
    geom_point(size=s1, shape=21, alpha=0.5) +
    xlab(paste0("PC1: ",percentVar[1],"% variance")) +
    ylab(paste0("PC2: ",percentVar[2],"% variance")) +
    scale_color_manual(name="Project",
                       breaks=levels(pc_data1[,"Project"]),
                       values=c(gdc_color , other_color)) 
  p1
}, x= tcga_fls1, y = percentVar_fls, SIMPLIFY = FALSE)


#  make GTEx pairwise PCA plots
gtex_res = mapply(function(x, y){
  pc_data1 = read.delim(x, header=T, stringsAsFactors=FALSE)

  pc_data1$Project[which(pc_data1$Project=="XENA_Toil")]="Xena/Toil"
  pc_data1$Project[which(pc_data1$Project=="MSKCC_Norm")]="MSKCC"
  pc_data1$Project[which(pc_data1$Project=="MSKCC_BATCH")]="MSKCC Batch"
  pc_data1$Project[which(pc_data1$Project=="GTEX-V6")]="GTEx"

  other_title = setdiff(unique(pc_data1$Project), "GTEx")
  other_color = as.character(gtex_col_lst[other_title])
  gtex_color = as.character(gtex_col_lst["GTEx"])
  pc_data1$Project = factor(pc_data1$Project,
                            levels= c("GTEx", other_title))
  percentVar = read.delim( y, header=FALSE, stringsAsFactors = FALSE)[,1]
  p1 = ggplot(pc_data1, aes(PC1, PC2, color=Project)) +
    geom_point(size=s1, shape=21, alpha=0.5) +
    xlab(paste0("PC1: ",percentVar[1],"% variance")) +
    ylab(paste0("PC2: ",percentVar[2],"% variance")) +
    scale_color_manual(name="Project",
                       breaks=levels(pc_data1[,"Project"]),
                       values=c(gtex_color , other_color)) 
  p1
}, x=gtex_fls1, y=gtex_percentVar_list, SIMPLIFY = FALSE)


theme_hb <- theme_bw(base_family="Helvetica") +
  theme(plot.title = element_text(hjust=0, vjust=0, 
                        lineheight=.8, face="bold", size=plot_title_size ),
  plot.margin=unit(chosen_margin,"cm"), 
  axis.text=element_text(size=axis_text_size),
  axis.title=element_text(size=axis_title_size),
  legend.text=element_text(size=legend_text_size),
  legend.key.height = unit(spacing, "cm"),
  legend.position = "bottom",
  legend.title=element_blank() )

tcga_p1 = tcga_res[[1]] + theme_hb + ggtitle("a ") + 
  guides( color = guide_legend(override.aes =
            list(alpha = 1, size=legend_pt_size, shape=16, ncol=3)))

tcga_p2 = tcga_res[[2]] + theme_hb + ggtitle("b ") + 
  guides( color = guide_legend(override.aes =
            list(alpha = 1, size=legend_pt_size, shape=16, ncol=3)))

tcga_p3 = tcga_res[[3]] + theme_hb + ggtitle("c ") + 
  guides( color = guide_legend(override.aes =
            list(alpha = 1, size=legend_pt_size, shape=16, ncol=3)))

tcga_p4 = tcga_res[[4]] + theme_hb + ggtitle("d ") + 
  guides( color = guide_legend(override.aes =
            list(alpha = 1, size=legend_pt_size, shape=16, ncol=3)))

tcga_p5 = tcga_res[[5]] + theme_hb + ggtitle("e ") + 
  guides( color = guide_legend(override.aes =
            list(alpha = 1, size=legend_pt_size, shape=16, ncol=3)))

gtex_p1 = gtex_res[[1]] + theme_hb + ggtitle("a ") + 
  guides( color = guide_legend(override.aes =
            list(alpha = 1, size=legend_pt_size, shape=16, ncol=3)))

gtex_p2 = gtex_res[[2]] + theme_hb + ggtitle("b ") + 
  guides( color = guide_legend(override.aes =
            list(alpha = 1, size=legend_pt_size, shape=16, ncol=3)))

gtex_p3 = gtex_res[[3]] + theme_hb + ggtitle("c ") + 
  guides( color = guide_legend(override.aes =
            list(alpha = 1, size=legend_pt_size, shape=16, ncol=3)))

gtex_p4 = gtex_res[[4]] + theme_hb + ggtitle("d ") + 
  guides( color = guide_legend(override.aes =
            list(alpha = 1, size=legend_pt_size, shape=16, ncol=3)))

tex <- textGrob("")

# arrange and make final figure.
pdf(file.path( results_dir, "pdf", "Supp_Fig1_pairwise_TCGA_pca_plots.pdf"), 
    width = 24, height =16)
tcga_plot <- list(tcga_p1, tcga_p2, tcga_p3, tcga_p4, tcga_p5, tex )
layout = rbind( c(1, 2, 3), c(4, 5, 6))
marrangeGrob(tcga_plot,  layout_matrix = layout,
                    top =textGrob("") )
dev.off()

pdf(file.path( results_dir, "pdf", "Supp_Fig2_pairwise_GTEx_pca_plots.pdf"), 
    width = 24, height =16)
gtex_plot <- list(gtex_p1, gtex_p2, gtex_p3, gtex_p4,  tex )
layout = rbind( c(1, 2, 3), c(4, 5, 5))
marrangeGrob(gtex_plot, layout_matrix = layout,
                    top= textGrob(""))
dev.off()
```

## Pairwise expression plots for NPM1, IL3RA, OVCA2 and CEBPA

```{r}
rm(list=ls())
suppressPackageStartupMessages({
  library(SummarizedExperiment)
  library(Hmisc)
  library(ggplot2)
  library(pheatmap)
  library(RColorBrewer)
  library(grid)
  library(gridExtra)
})

# folder where S3BUCKET data and github directory are stored. eg: ~/Downloads
bigdir = dirname(getwd())
# github directory eg: ~/Downloads/UncertaintyRNA
git_dir = file.path(bigdir,  "UncertaintyRNA")
# S3 bucket directory eg: ~/Downloads/OriginalTCGAGTExData
s3_dir = file.path(bigdir,  "OriginalTCGAGTExData")

# when you run our RMD files, all results will be stored here. 
# This will essentially remake the "data" subfolder from github repo.
# eg:~/Downloads/data
results_dir = file.path(bigdir, "data")


if(!file.exists(file.path( s3_dir, "SE_objects"))){
  stop("Please go through vignette 3 & 4 to make SE objects or download from S3 bucket")
}

tcga_gdc <- get(load( file.path( s3_dir, "SE_objects","tcga_gdc_log2_TPM.RData")))
tcga_mskcc_norm <- get(load( file.path( s3_dir, "SE_objects", "tcga_mskcc_norm_log2_TPM.RData")))
tcga_mskcc_batch <- get(load( file.path( s3_dir, "SE_objects", "tcga_mskcc_batch_log2_TPM.RData")))
tcga_recount2 <- get(load( file.path( s3_dir, "SE_objects", "tcga_recount2_log2_TPM.RData")))
tcga_xena <- get(load( file.path( s3_dir, "SE_objects", "tcga_xena_log2_TPM.RData")))
tcga_piccolo <- get(load( file.path( s3_dir, "SE_objects","tcga_piccolo_log2_TPM.RData")))

geneName = rownames(tcga_gdc)

gdc_mat = assay(tcga_gdc)
mskcc_norm_mat=assay(tcga_mskcc_norm)
mskcc_batch_mat=assay(tcga_mskcc_batch)
piccolo_mat=assay(tcga_piccolo)
recount2_mat=assay(tcga_recount2)
xena_mat= assay(tcga_xena)

m = 5 # pairwise max of 2 vectors.
lf = 2 # pairwise logfold change difference between 2 vectors

plot_title_size =20
axis_text_size = 20

myFun  = function(x1, y1, plot_title){
  df = data.frame(cbind( x=x1, y=y1))
  sel = which(abs(x1- y1) > lf & pmax(x1, y1) > m)
  color1 = rep("black", 4800)
  color1[sel] = "red"
  p0 = ggplot(df, aes(x=x, y=y)) +
    geom_point(colour=color1) +
    theme_bw(base_family="Helvetica") +
    xlab("") +
    ylab("")+
    ggtitle(plot_title) +
    theme(plot.title = element_text(lineheight=.8, size=plot_title_size),
          axis.text=element_text(size=axis_text_size), 
          axis.title=element_text(size=axis_text_size), 
          plot.margin=unit(c(1,0.5,0.5,0.5),"cm")) # margins:top,right,bottom,left

  p0
}

selGenes <- c("CEBPA","NPM1") # "IL3RA","OVCA2", "U2AF1")

plots <- lapply( selGenes, function(gene){
  idx = grep(paste0("^",gene,"$"), rownames(gdc_mat))
  temp_gdc=gdc_mat[idx, ]
  temp_piccolo=piccolo_mat[idx, ]
  temp_mskcc_norm=mskcc_norm_mat[idx, ]
  temp_recount2=recount2_mat[idx, ]
  temp_xena=xena_mat[idx, ]
  
  if(gene == "U2AF1"){
    u2af1_lst =list(temp_gdc =temp_gdc, 
                    temp_piccolo =temp_piccolo, 
                    temp_mskcc_norm = temp_mskcc_norm, 
                    temp_recount2 = temp_recount2, 
                    temp_xena = temp_xena) 
    save(u2af1_lst, file="data/discordant/u2af1_tcga_expression.RData")
  }
  
  p1 = myFun(x=temp_gdc, y=temp_xena, plot_title = "Xena/Toil vs\n GDC")
  p1 = p1 + ylab(paste0(gene, " (log2(TPM+0.001))")) 
  
  p2 = myFun(x=temp_gdc, y=temp_recount2, plot_title = "Recount2 vs\n GDC")
  p3 = myFun(x=temp_gdc, y=temp_mskcc_norm, plot_title = "MSKCC \nvs GDC")
  p4 = myFun(x=temp_gdc, y=temp_piccolo, plot_title = "Piccolo vs\n GDC")
  p5 = myFun(x=temp_xena, y=temp_recount2, plot_title = "Recount2 vs\n Xena/Toil")
  
  p6 = myFun(x=temp_xena, y=temp_piccolo, plot_title = "Piccolo vs\n Xena/Toil")
  p6 = p6 + ylab(paste0(gene, " (log2(TPM+0.001))"))
  
  p7 = myFun(x=temp_xena, y=temp_mskcc_norm, plot_title = "MSKCC \nvs Xena/Toil")
  p8 = myFun(x=temp_recount2, y=temp_piccolo, plot_title = "Piccolo vs\n Recount2")
  p9 = myFun(x=temp_recount2, y=temp_mskcc_norm, plot_title = "MSKCC \nvs Recount2")
  p10 = myFun(x=temp_mskcc_norm, y=temp_piccolo, plot_title = "Piccolo vs\n MSKCC")
  
 list(p1, p2, p3, p4, p5, p6, p7, p8, p9 , p10)
})

allplots = list(plots[[1]], plots[[2]])
allplots = do.call( c, allplots)

pdf(file.path(results_dir, "pdf", "Supp_Fig6_Pairwise_expression.pdf"), 
    width = 24, height = 16)

grid.arrange(grobs = allplots ,
             layout_matrix =rbind( c(1, 2, 3, 4, 5),
                                   c(6, 7, 8, 9, 10 ),
                                   c(11, 12, 13, 14, 15), 
                                   c(16, 17, 18, 19, 20 )), 
             heights=unit(c(4,4, 4,4), c("in", "in", "in", "in")))
dev.off()

```

## Pairwise scatter plots for Discordant genes in TCGA samples

```{r}
rm(list=ls())
suppressPackageStartupMessages({
  library(SummarizedExperiment)
  library(Hmisc)
  library(ggplot2)
  library(grid)
  library(gridExtra)
})

# folder where S3BUCKET data and github directory are stored. eg: ~/Downloads
bigdir = dirname(getwd())
# github directory eg: ~/Downloads/UncertaintyRNA
git_dir = file.path(bigdir,  "UncertaintyRNA")
# S3 bucket directory eg: ~/Downloads/OriginalTCGAGTExData
s3_dir = file.path(bigdir,  "OriginalTCGAGTExData")

# when you run our RMD files, all results will be stored here. 
# This will essentially remake the "data" subfolder from github repo.
# eg:~/Downloads/data
results_dir = file.path(bigdir, "data")


if(!file.exists(file.path( s3_dir, "SE_objects"))){
  stop("Please go through vignette 3 & 4 to make SE objects or download from S3 bucket")
}

tcga_gdc <- get(load( file.path( s3_dir, "SE_objects","tcga_gdc_log2_TPM.RData")))
tcga_mskcc_norm <- get(load( file.path( s3_dir, "SE_objects", "tcga_mskcc_norm_log2_TPM.RData")))
tcga_mskcc_batch <- get(load( file.path( s3_dir, "SE_objects", "tcga_mskcc_batch_log2_TPM.RData")))
tcga_recount2 <- get(load( file.path( s3_dir, "SE_objects", "tcga_recount2_log2_TPM.RData")))
tcga_xena <- get(load( file.path( s3_dir, "SE_objects", "tcga_xena_log2_TPM.RData")))
tcga_piccolo <- get(load( file.path( s3_dir, "SE_objects","tcga_piccolo_log2_TPM.RData")))

geneName = rownames(tcga_gdc)

gdc_mat = assay(tcga_gdc)
mskcc_norm_mat=assay(tcga_mskcc_norm)
mskcc_batch_mat=assay(tcga_mskcc_batch)
piccolo_mat=assay(tcga_piccolo)
recount2_mat=assay(tcga_recount2)
xena_mat= assay(tcga_xena)

m = 5 # pairwise max of 2 vectors.
lf = 2; fold_no = "4fold"

selGenes <- read.delim(file.path(git_dir, "data", "discordant", 
  "TCGA_bad_genes_samples_varying_fc_across_pipelines.txt"), header=TRUE, 
  stringsAsFactors = FALSE)[,1]


plot_title_size =15
axis_text_size = 15

myFun  = function(x1, y1, plot_title){
  df = data.frame(cbind( x=x1, y=y1))
  sel = which(abs(x1- y1) > lf & pmax(x1, y1) > m)
  color1 = rep("black", 4800)
  color1[sel] = "red"
  p0 = ggplot(df, aes(x=x, y=y)) +
    geom_point(colour=color1) +
    theme_bw(base_family="Helvetica") +
    xlab("") +
    ylab("")+
    ggtitle(plot_title) +
    scale_y_continuous(breaks=seq(2,10,2)) +
    theme(plot.title = element_text(lineheight=.8, size=plot_title_size),
          axis.text=element_text(size=axis_text_size),
          axis.title = element_text(size=axis_text_size))
  
  p0
}

pdf(file.path(results_dir, "pdf", "Supp_Fig_TCGA_discordant_genes_examples.pdf"),
   width = 12, height = 7)
plots <- lapply( selGenes, function(gene){
  idx = grep(paste0("^",gene,"$"), rownames(gdc_mat))
  temp_gdc=gdc_mat[idx, ]
  temp_piccolo=piccolo_mat[idx, ]
  temp_mskcc_norm=mskcc_norm_mat[idx, ]
  temp_recount2=recount2_mat[idx, ]
  temp_xena=xena_mat[idx, ]
  
  p1 = myFun(x=temp_gdc, y=temp_xena, plot_title = "Xena/Toil vs\n GDC")
  p1 = p1 + ylab(paste0(gene, " (log2(TPM+0.001))")) 
  
  p2 = myFun(x=temp_gdc, y=temp_recount2, plot_title = "Recount2 vs\n GDC")
  p3 = myFun(x=temp_gdc, y=temp_mskcc_norm, plot_title = "MSKCC \nvs GDC")
  p4 = myFun(x=temp_gdc, y=temp_piccolo, plot_title = "Piccolo vs\n GDC")
  p5 = myFun(x=temp_xena, y=temp_recount2, plot_title = "Recount2 vs\n Xena/Toil")
  
  p6 = myFun(x=temp_xena, y=temp_piccolo, plot_title = "Piccolo vs\n Xena/Toil")
  p6 = p6 + ylab(paste0(gene, " (log2(TPM+0.001))"))
  
  p7 = myFun(x=temp_xena, y=temp_mskcc_norm, plot_title = "MSKCC \nvs Xena/Toil")
  p8 = myFun(x=temp_recount2, y=temp_piccolo, plot_title = "Piccolo vs\n Recount2")
  p9 = myFun(x=temp_recount2, y=temp_mskcc_norm, plot_title = "MSKCC \nvs Recount2")
  p10 = myFun(x=temp_mskcc_norm, y=temp_piccolo, plot_title = "Piccolo vs\n MSKCC")
  
  allplots <- list(p1, p2, p3, p4, p5, p6, p7, p8, p9 , p10)
 
  layout = rbind( c(1,2,3,4,5), c(6,7,8,9,10))
  ml <- marrangeGrob(allplots, layout_matrix = layout,
                     top= textGrob(""))
  print(ml)
})
dev.off()


```

## Pairwise scatter plots for Discordant genes in GTEX samples

```{r}
rm(list=ls())
suppressPackageStartupMessages({
  library(SummarizedExperiment)
  library(ggplot2)
  library(grid)
  library(gridExtra)
})

# folder where S3BUCKET data and github directory are stored. eg: ~/Downloads
bigdir = dirname(getwd())
# github directory eg: ~/Downloads/UncertaintyRNA
git_dir = file.path(bigdir,  "UncertaintyRNA")
# S3 bucket directory eg: ~/Downloads/OriginalTCGAGTExData
s3_dir = file.path(bigdir,  "OriginalTCGAGTExData")

# when you run our RMD files, all results will be stored here. 
# This will essentially remake the "data" subfolder from github repo.
# eg:~/Downloads/data
results_dir = file.path(bigdir, "data")

gtex_v6 <- get(load( file.path( s3_dir, "SE_objects","gtex_v6_log2_TPM.RData")))
gtex_mskcc_norm <- get(load( file.path( s3_dir, "SE_objects","gtex_mskcc_norm_log2_TPM.RData")))
gtex_recount2 <- get(load( file.path( s3_dir, "SE_objects", "gtex_recount2_log2_TPM.RData")))
gtex_xena <- get(load( file.path( s3_dir, "SE_objects","gtex_xena_log2_TPM.RData")))

gtex_v6_mat = assay(gtex_v6)
mskcc_norm_mat=assay(gtex_mskcc_norm)
recount2_mat=assay(gtex_recount2)
xena_mat= assay(gtex_xena)

selGenes <- read.delim(file.path(git_dir, "data", "discordant", 
  "GTEX_bad_genes_samples_varying_fc_across_pipelines.txt"), header=TRUE, 
  stringsAsFactors = FALSE)[,1]

m = 5 # pairwise max of 2 vectors.
lf = 2; fold_no = "4fold"

#  make pairwise plots to check that these genes are actually discordant
plot_title_size =15
axis_text_size = 15

myFun  = function(x1, y1, plot_title){
  df = data.frame(cbind( x=x1, y=y1))
  sel = which(abs(x1- y1) > lf & pmax(x1, y1) > m)
  color1 = rep("black", 1890)
  color1[sel] = "red"
  p0 = ggplot(df, aes(x=x, y=y)) +
    geom_point(colour=color1) +
    theme_bw(base_family="Helvetica") +
    xlab("") +
    ylab("")+
    ggtitle(plot_title) +
    scale_y_continuous(breaks=seq(2,10,2)) +
    theme(plot.title = element_text(lineheight=.8, size=plot_title_size),
          axis.text=element_text(size=axis_text_size),
          axis.title = element_text(size=axis_text_size))
  
  p0
}

pdf(file.path(results_dir, "pdf",
  "Supp_Fig_GTEX_discordant_genes_examples.pdf"), width = 12, height = 7)
plots <- lapply( selGenes, function(gene){
  idx = grep(paste0("^",gene,"$"), rownames(gtex_v6_mat))
  temp_gtex=gtex_v6_mat[idx, ]
  temp_mskcc_norm=mskcc_norm_mat[idx, ]
  temp_recount2=recount2_mat[idx, ]
  temp_xena=xena_mat[idx, ]
  
  p1 = myFun(x=temp_gtex, y=temp_xena, plot_title = "Xena/Toil vs\n GTEX")
  p1 = p1 + ylab(paste0(gene, " (log2(TPM+0.001))")) 
  
  p2 = myFun(x=temp_gtex, y=temp_recount2, plot_title = "Recount2 vs\n GTEX")
  
  p3 = myFun(x=temp_gtex, y=temp_mskcc_norm, plot_title = "MSKCC \nvs GTEX")
  
  p5 = myFun(x=temp_xena, y=temp_recount2, plot_title = "Recount2 vs\n Xena/Toil")
  
  p7 = myFun(x=temp_xena, y=temp_mskcc_norm, plot_title = "MSKCC \nvs Xena/Toil")
  
  p9 = myFun(x=temp_recount2, y=temp_mskcc_norm, plot_title = "MSKCC \nvs Recount2")
  p9 = p9 + ylab(paste0(gene, " (log2(TPM+0.001))"))
  
  allplots <- list(p1, p2, p3, p5,  p7,  p9 )
  
  layout = rbind( c(1, 2, 3,4,5), c(6,7,8,9,10))
  ml <- marrangeGrob(allplots, layout_matrix = layout,
                     top= textGrob(""))
  print(ml)
})
dev.off()


```

## Correlations between each source of TCGA and GTEx Data

```{r}
rm(list=ls())

suppressPackageStartupMessages({
  library(ggplot2)
  library(grid)
  library(gridExtra)
  library(beeswarm)
})

s1 = 3 # size for points in PCA plot
legend_pt_size =4
plot_title_size = 25
axis_text_size = 25
axis_title_size=25
legend_text_size=20
spacing=0.3
chosen_margin = c(0.5,1,0.5,1)# margins:top,right,bottom,left
text_size =10

# folder where S3BUCKET data and github directory are stored. eg: ~/Downloads
bigdir = dirname(getwd())
# github directory eg: ~/Downloads/UncertaintyRNA
git_dir = file.path(bigdir,  "UncertaintyRNA")
# S3 bucket directory eg: ~/Downloads/OriginalTCGAGTExData
s3_dir = file.path(bigdir,  "OriginalTCGAGTExData")

# when you run our RMD files, all results will be stored here. 
# This will essentially remake the "data" subfolder from github repo.
# eg:~/Downloads/data
results_dir = file.path(bigdir, "data")


if(!file.exists( file.path( git_dir, "data", "tables"))){
  stop("Please clone from github")
}
if(!file.exists( file.path( results_dir))){
   system(paste0("mkdir ", results_dir))
}
if(!file.exists( file.path( results_dir, "pdf"))){
   system(paste0("mkdir ", file.path(results_dir, "pdf")))
}


tcga_result_file = file.path( git_dir, "data", "tables", 
                "Supp_Table_TCGA_Correlations_all_6_Datasets_log2_TPM.txt")

gtex_result_file = file.path( git_dir, "data", "tables", 
    "Supp_Table_GTEX_Correlations_all_6_Datasets_log2_TPM.txt")

protein_result_file =file.path( git_dir, "data", "tables", 
          "Supp_Table_TCGA_Protein_MRNA_Correlations_all_6_Datasets.txt")

tcga_mat = read.delim(tcga_result_file,
                      header=TRUE,  stringsAsFactors = FALSE, row.names=1)
gtex_mat = read.delim(gtex_result_file,
                      header=TRUE,  stringsAsFactors = FALSE, row.names=1)
mat = read.delim(protein_result_file,
                 header=TRUE, stringsAsFactors = FALSE )


tcga_data = rbind(
  cbind( Value= tcga_mat[,"corr_XENA_Toil"], Group= rep("Xena/Toil", nrow(tcga_mat))) ,
  cbind( Value= tcga_mat[,"corr_Piccolo_Lab"], Group= rep("Piccolo", nrow(tcga_mat))) ,
  cbind( Value= tcga_mat[,"corr_Recount2"], Group= rep("Recount2", nrow(tcga_mat))) ,
  cbind( Value= tcga_mat[,"corr_MSKCC_norm"], Group= rep("MSKCC", nrow(tcga_mat))) ,
  cbind( Value= tcga_mat[,"corr_MSKCC_BATCH"], Group= rep("MSKCC Batch", nrow(tcga_mat))) )
group_levels=c("Xena/Toil", "Piccolo","Recount2", 
               "MSKCC","MSKCC Batch")
group_colors= c("orange", "#E76BF3","grey45", "darkblue", "#00B0F6" )

tcga_data = as.data.frame(tcga_data, stringsAsFactors=FALSE)
tcga_data[,1] = as.numeric(tcga_data[,1])
tcga_data =tcga_data[ complete.cases(tcga_data[,1]), ]
tcga_data = tcga_data[-c(which(tcga_data[,1] == -1)), ]

tcga_data[,2] = factor(tcga_data[,2],levels = group_levels)

dodge <- position_dodge(width = 0.4)
tcga_cor= ggplot(tcga_data, aes(x = Group, y = Value, color=Group )) +
  geom_violin(position=dodge, fill= "gray60") +
  xlab("") + ylab("Pearson Correlation") +
  ggtitle("a")  +
  geom_boxplot(width=0.1, outlier.color = NA, position=dodge, alpha=0.5) +
  theme_bw(base_family="Helvetica") +
  theme(plot.title = element_text(lineheight=.8, face="bold", size=plot_title_size),
        plot.margin=unit(chosen_margin,"cm"), 
        axis.text.x = element_text(angle = 45, hjust = 1, colour = "black"),
        axis.text=element_text(size=axis_text_size),
        axis.title=element_text(size=axis_title_size),
        legend.text=element_text(size=legend_text_size),
        legend.key.height = unit(spacing, "cm"),
        #legend.position = "bottom",
        legend.title=element_blank()) +
    scale_color_manual(name = "Data Source",
                     labels = group_levels,
                     values = group_colors) +
  guides(fill=FALSE) +
  guides(color=FALSE) +
  annotate("text", label = length(which(tcga_mat[,"corr_XENA_Toil"] < 0.8)) ,
           x = 0.8, y = 0.75, size = text_size, colour = "red") +
  annotate("text", label = length(which(tcga_mat[,"corr_Piccolo_Lab"] < 0.8))  ,
           x = 1.8, y = 0.75, size = text_size, colour = "red") +
  annotate("text", label = length(which(tcga_mat[,"corr_Recount2"] < 0.8))  ,
           x = 2.8, y = 0.75, size = text_size, colour = "red") +
  annotate("text", label = length(which(tcga_mat[,"corr_MSKCC_norm"] < 0.8))  ,
           x = 3.8, y = 0.75, size = text_size, colour = "red") +
  annotate("text", label = length(which(tcga_mat[,"corr_MSKCC_BATCH"] < 0.8)) ,
           x = 4.8, y = 0.75, size = text_size, colour = "red") +
  geom_hline(aes(yintercept = 0.8), linetype="dashed", colour="red")


gtex_data = rbind(
  cbind( Value= gtex_mat[,"corr_XENA_Toil"], Group= rep("Xena/Toil", nrow(gtex_mat))) ,
  cbind( Value= gtex_mat[,"corr_Recount2"], Group= rep("Recount2", nrow(gtex_mat))) ,
  cbind( Value= gtex_mat[, "corr_MSKCC_Norm"], Group= rep("MSKCC", nrow(gtex_mat))) ,
  cbind( Value= gtex_mat[,"corr_MSKCC_BATCH"], Group= rep("MSKCC Batch", nrow(gtex_mat))) )

group_levels2=c("Xena/Toil", "Recount2","MSKCC", "MSKCC Batch")
group_colors2= c("orange","grey45", "darkblue", "#00B0F6")

gtex_data = as.data.frame(gtex_data, stringsAsFactors=FALSE)
gtex_data[,1] = as.numeric(gtex_data[,1])
gtex_data =gtex_data[ complete.cases(gtex_data[,1]), ]

gtex_data[,2] = factor(gtex_data[,2], levels =group_levels2 )


dodge <- position_dodge(width = 0.4)
gtex_cor=ggplot(gtex_data, aes(x = Group, y = Value, fill=Group)) +
  geom_violin(position=dodge) +
  xlab("") + ylab("Pearson Correlation") +
  ggtitle("b")  +
  geom_boxplot(width=.1, outlier.color = NA, position=dodge, fill="white") +
  theme_bw(base_family="Helvetica") +
  theme(plot.title = element_text(lineheight=.8, face="bold", size=plot_title_size),
        plot.margin=unit(chosen_margin,"cm"), 
        axis.text.x = element_text(angle = 45, hjust = 1, colour = "black"),
        axis.text=element_text(size=axis_text_size),
        axis.title=element_text(size=axis_title_size),
        legend.text=element_text(size=legend_text_size),
        legend.key.height = unit(spacing, "cm"),
        legend.title=element_blank()) +
  scale_fill_manual(name = "Data Source",
                    labels = group_levels2,
                    values = group_colors2) +
  guides(fill=FALSE) +
  annotate("text", label = length(which(gtex_mat[,"corr_XENA_Toil"] < 0.8))  ,
           x = 0.7, y = 0.75, size = text_size, colour = "red") +
  annotate("text", label = length(which(gtex_mat[,"corr_Recount2"] < 0.8))  ,
           x = 1.7, y = 0.75, size = text_size, colour = "red") +
  annotate("text", label = length(which(gtex_mat[,"corr_MSKCC_Norm"] < 0.8))  ,
           x = 2.7, y = 0.75, size = text_size, colour = "red") +
  annotate("text", label = length(which(tcga_mat[,"corr_MSKCC_BATCH"] < 0.8)) ,
           x = 3.7, y = 0.75, size = text_size, colour = "red") +
  geom_hline(aes(yintercept = 0.8), linetype="dashed", colour="red")



prot_data = rbind(
  cbind( Value= mat[,"corr_protein_GDC"], Group= rep("GDC", nrow(mat))) ,
  cbind( Value= mat[,"corr_protein_PiccoloLab"], Group= rep("Piccolo", nrow(mat))) ,
  cbind( Value= mat[,"corr_protein_mskcc_BATCH"], Group= rep("MSKCC Batch", nrow(mat))) ,
  cbind( Value= mat[,"corr_protein_mskcc_Norm"], Group= rep("MSKCC", nrow(mat))) ,
  cbind( Value= mat[,"corr_protein_Recount2"], Group= rep("Recount2", nrow(mat))),
  cbind( Value= mat[,"corr_protein_Xena_Toil"], Group= rep("Xena/Toil", nrow(mat)))   )

group_levels3=c("GDC", "Xena/Toil", "Piccolo", "Recount2","MSKCC", "MSKCC Batch")
group_colors3= c("#A3A500",  "orange","#E76BF3", "grey45", "darkblue", "#00B0F6")
   
prot_data = as.data.frame(prot_data, stringsAsFactors=FALSE)
prot_data[,1] = as.numeric(prot_data[,1])
prot_data[,2] = factor(prot_data[,2],levels = group_levels3)

beeswarm = beeswarm(Value ~Group, data = prot_data, method = "swarm")
colnames(beeswarm) = gsub("x.orig", "DataSource", colnames(beeswarm))
beeswarm[,"DataSource"] = factor(beeswarm[,"DataSource"],levels = group_levels3)

dodge <- position_dodge(width = 0.4)
prot_cor =   ggplot(beeswarm, aes(x, y, colour = DataSource)) +
  geom_boxplot(aes(x, y, group = DataSource),
               width=.1, outlier.color = NA, position=dodge, fill="white") +
  geom_point( aes(colour=DataSource), shape=16) +
  xlab("") + ylab("Spearman Correlation") +
  ggtitle("c")  +
  scale_color_manual(name = "Data Source",
                     labels = group_levels3,
                     values = group_colors3) +
  theme_bw(base_family="Helvetica") +
  theme(plot.title = element_text(lineheight=.8, face="bold", size=plot_title_size),
        plot.margin=unit(chosen_margin,"cm"),
        axis.text=element_text(size=axis_text_size),
        axis.title=element_text(size=axis_title_size),
        legend.text=element_text(size=legend_text_size),
        legend.key.height = unit(spacing, "cm"),
        axis.text.x = element_text(angle = 45, hjust = 1, colour = "black"),
        legend.title=element_blank(), 
        legend.position="none")  +
  scale_x_continuous(breaks = c(1:6),
                     labels = group_levels3)

pdf(file.path( results_dir, "pdf", "Supp_Fig7_correlations.pdf"), 
    width =24, height =12)
layout <- rbind( c(1, 2, 3))
nat3= list(tcga_cor, gtex_cor, prot_cor)
marrangeGrob(nat3, layout_matrix = layout, top="")
dev.off()

```
