---
title: "Pathway Analysis for cancer types"
author: "Sonali Arora, Hamid Bolouri"
date: "November 18, 2019"
output: 
  html_document:
    toc: true
    theme: united
---



## Pathway Analysis ( Over-representation analysis)

In the previous vignette, we found differentially expressed genes (DEGs) using DESeq2. 
In the below code chunk, we take those differentially expressed genes and perform a pathway
over-representation analysis for REACTOME pathways using R/Bioconductor package clusterProfileR.

```{r eval = FALSE}

rm(list=ls())
suppressPackageStartupMessages({
  library(DOSE)
  library(clusterProfiler)
  library(ReactomePA)
  library(org.Hs.eg.db)
  library(DESeq2)
  library(edgeR)
  library(RColorBrewer)
  library(pheatmap)
  library(ggplot2)
})

getEntrezIdFromSymbol <- function(gene_symbol){
  mapIds(org.Hs.eg.db,
         keys=gene_symbol,
         column="ENTREZID",
         keytype="SYMBOL",
         multiVals="first")
}

getSymbolFromEntrezId <- function(gene_id){
  mapIds(org.Hs.eg.db,
         keys=gene_id,
         column="SYMBOL",
         keytype="ENTREZID",
         multiVals="first")
}

myEnrichmentFunction = function(fls, resdir){
  
  gene_list = read.delim(fls, header=T, stringsAsFactors = FALSE)[,1]
  title = gsub(".txt", "", basename(fls))
  
  up_ids = getEntrezIdFromSymbol(unlist(gene_list))
  reactome = enrichPathway(up_ids, organism = "human", pvalueCutoff = 0.05,
                           pAdjustMethod = "BH",   minGSSize = 10,
                           maxGSSize = 500, readable = FALSE)
  
  message(" No of genes: ", length(unique(gene_list)))
  message("Reactome for up-reg genes=", nrow(reactome))
 
  if(nrow(reactome)!=0){
    r1 = as.data.frame(reactome)
    gene_name= lapply(r1$geneID, function(y){
      y = unlist(strsplit(y, "/"))
      paste0(getSymbolFromEntrezId(y), collapse="/")
    })
    r1$gene_name = unlist(gene_name)
    
    write.table(r1,
                file.path(resdir, paste0(title,"_Reactome_Pathway.txt")),
                sep="\t", quote=FALSE, row.names=FALSE, col.names=TRUE)
    
  }
  rm(reactome)
}


# folder where S3BUCKET data and github directory are stored. eg: ~/Downloads
bigdir = dirname(getwd())
# github directory eg: ~/Downloads/UncertaintyRNA
git_dir = file.path(bigdir,  "UncertaintyRNA")
# S3 bucket directory eg: ~/Downloads/OriginalTCGAGTExData
s3_dir = file.path(bigdir,  "OriginalTCGAGTExData", "raw_counts")

# when you run our RMD files, all results will be stored here. 
# This will essentially remake the "data" subfolder from github repo.
# eg:~/Downloads/data/de_analysis
resdir = file.path(bigdir, "data", "de_analysis")


down_fls = list.files(pattern = "_down_reg_genes.txt", path = resdir, full.names=TRUE, recursive=TRUE)
up_fls = list.files(pattern = "_up_reg_genes.txt", path = resdir, full.names=TRUE, recursive = TRUE)

lapply(up_fls, function(x){
  myEnrichmentFunction(x, resdir=  file.path(resdir, "pathway_analysis"))
})

lapply(down_fls, function(x){
  message(x)
  myEnrichmentFunction(x, resdir=  file.path(resdir, "pathway_analysis"))
})

```


## Combined dotplots for each cancer comparison

In the code section below, we create dotplots to visualize pathways over-represented in each cancer type 
across all pipelines. This is included as a three panel supplemental figure in our manuscript.


```{r eval=FALSE}
rm(list=ls())
library(ggplot2)


bigdir = dirname(getwd())
# github directory eg: ~/Downloads/UncertaintyRNA
git_dir = file.path(bigdir,  "UncertaintyRNA")
# S3 bucket directory eg: ~/Downloads/OriginalTCGAGTExData
s3_dir = file.path(bigdir,  "OriginalTCGAGTExData", "raw_counts")

# when you run our RMD files, all results will be stored here. 
# This will essentially remake the "data" subfolder from github repo.
# eg:~/Downloads/data/de_analysis
resdir = file.path(bigdir, "data", "de_analysis", "pathway_analysis")

fls = list.files(pattern ="up_reg_genes_Reactome_Pathway.txt", recursive=T, 
                 full.names=TRUE, path = resdir)

parse_ratio <- function(ratio) {
  gsize <- as.numeric(sub("/\\d+$", "", as.character(ratio)))
  gcsize <- as.numeric(sub("^\\d+/", "", as.character(ratio)))
  return(gsize/gcsize)
}

tag_lst = c("HER2_enriched_vs_basal_like_", 
            "kirp_vs_kirc_", 
            "lusc_vs_luad_")
file_ending = "_deseq2_up_reg_genes_Reactome_Pathway.txt"

lapply(tag_lst, function(tag){
  message(tag)
    small_fls = grep(tag, fls, value=T)
    ppl_lst = c("GDC", "MSKCC", "Piccolo", "Recount2", "Xena/Toil")
    
    bigdf = lapply(small_fls, function(x){
      df  = read.delim(x, header=T, stringsAsFactors = FALSE)
      pipeline = gsub(tag, "", basename(x))
      pipeline = gsub(file_ending, "", pipeline)
      df$pipeline = rep(pipeline, nrow(df))
      df$GeneRatio <- parse_ratio(df$GeneRatio)
      df[, c("Description", "GeneRatio", "p.adjust", "pipeline")]
    })
    
    bigdf = do.call(rbind, bigdf)
    
    bigdf$pipeline = gsub("gdc", "GDC", bigdf$pipeline)
    bigdf$pipeline = gsub("mskcc", "MSKCC", bigdf$pipeline)
    bigdf$pipeline = gsub("XENA", "Xena/Toil", bigdf$pipeline)
    bigdf$pipeline = gsub("RECOUNT2", "Recount2", bigdf$pipeline)
    
    # edit long text for one pathway to ensure it fits on pdf.
    long_text = "Regulation of Insulin-like Growth Factor (IGF) transport and uptake by Insulin-like Growth Factor Binding Proteins (IGFBPs)"
    small_text = "Regulation of Insulin-like Growth Factor (IGF) transport"
    long1 = which(bigdf$Description ==long_text)
    if(length(long1)!=0){
      bigdf$Description[long1]= small_text
    }
    
    test = as.data.frame(table(bigdf$Description))
    test = test[order(test[,2], decreasing=T), ]
    
    # print only 25 pathways.
    n = 25
    if(nrow(test)< 25){
      n = nrow(test)
    }
    test = tail(test, n)
    test[,1] = as.character(test[,1])
    idx = sapply(test[,1], function(x) which(bigdf$Description==x))
    smalldf = bigdf[ unlist(idx), ]
    
    smalldf$Description = factor(smalldf$Description, levels = test[,1] )
    p2 = ggplot(smalldf, aes(x=pipeline, y=Description, size=GeneRatio, color=p.adjust)) +
      geom_point() + scale_color_gradient(low="red", high="blue") + ylab("") +
      theme_bw()+  xlab("") + theme(axis.text.x = element_text(angle = 45, hjust = 1))
    
    pdf(file.path(resdir, paste0("Combined_dotplot_", tag, gsub(".txt", ".pdf", file_ending))), width =7, height = 6)
    print(p2)
    dev.off()
})

```


## Creating files as input for GSEA

In addition to over-represented pathways ( using DEGs), we also decided to do a Gene Set Enrichment Analysis (GSEA)
using log2(tpm+1) counts from each pipeline. 
In the below section, we create input files for GSEA. 

### Breast

```{r eval = FALSE}
rm(list=ls())
library(SummarizedExperiment)
s3_dir = "~/HollandLabShared/Sonali/paper_with_Hamid_Siobhan/Rstudio/OriginalTCGAGTExData/"

# folder where S3BUCKET data and github directory are stored. eg: ~/Downloads
bigdir = dirname(getwd())
# github directory eg: ~/Downloads/UncertaintyRNA
git_dir = file.path(bigdir,  "UncertaintyRNA")
# S3 bucket directory eg: ~/Downloads/OriginalTCGAGTExData
s3_dir = file.path(bigdir,  "OriginalTCGAGTExData")

# when you run our RMD files, all results will be stored here. 
# This will essentially remake the "data" subfolder from github repo.
# eg:~/Downloads/data/de_analysis
resdir = file.path(bigdir, "data", "gsea_analysis")
gsea_dir = file.path(bigdir, "data", "gsea_analysis")

tcga_gdc <- get(load( file.path( s3_dir, "SE_objects","tcga_gdc_log2_TPM.RData")))
tcga_mskcc_norm <- get(load( file.path( s3_dir, "SE_objects", "tcga_mskcc_norm_log2_TPM.RData")))
tcga_mskcc_batch <- get(load( file.path( s3_dir, "SE_objects", "tcga_mskcc_batch_log2_TPM.RData")))
tcga_recount2 <- get(load( file.path( s3_dir, "SE_objects", "tcga_recount2_log2_TPM.RData")))
tcga_xena <- get(load( file.path( s3_dir, "SE_objects", "tcga_xena_log2_TPM.RData")))
tcga_piccolo <- get(load( file.path( s3_dir, "SE_objects","tcga_piccolo_log2_TPM.RData")))

# sanity check - check rows and columns are identical for easier subsetting ahead

identical( rownames(tcga_gdc), rownames(tcga_mskcc_batch))
identical( rownames(tcga_gdc), rownames(tcga_mskcc_norm))
identical( rownames(tcga_gdc), rownames(tcga_piccolo))
identical( rownames(tcga_gdc), rownames(tcga_recount2))
identical( rownames(tcga_gdc), rownames(tcga_xena))

identical( colnames(tcga_gdc), colnames(tcga_mskcc_batch))
identical( colnames(tcga_gdc), colnames(tcga_mskcc_norm))
identical( colnames(tcga_gdc), colnames(tcga_piccolo))
identical( colnames(tcga_gdc), colnames(tcga_recount2))
identical( colnames(tcga_gdc), colnames(tcga_xena))

tag = "Breast"

goi= read.delim(file.path(s3_dir, "de_goi.txt"), header=FALSE, stringsAsFactors = FALSE )[,1]
basal_like_samples= read.delim(file.path(s3_dir,"brca_basal_like_samples.txt"), header=TRUE, stringsAsFactors = FALSE )[,1]
HER2_enriched_samples= read.delim(file.path(s3_dir,"brca_HER2_enriched_samples.txt"), header=TRUE, stringsAsFactors = FALSE )[,1]

idx1  = match( basal_like_samples, substr(colnames(tcga_gdc), 1, 12))
idx2 = match(HER2_enriched_samples, substr(colnames(tcga_gdc), 1, 12))
row_idx = match(goi, rownames(tcga_gdc))

table(is.na(idx1))
table(is.na(idx2))
table(is.na(row_idx))

her2_gdc = assay(tcga_gdc[goi, c(idx1, idx2)])
her2_xena = assay(tcga_xena[goi, c(idx1, idx2)])
her2_recount2 = assay(tcga_recount2[goi, c(idx1, idx2)])
her2_piccolo = assay(tcga_piccolo[ goi, c(idx1, idx2)])
her2_mskcc_norm = assay(tcga_mskcc_norm[goi, c(idx1,idx2)])
her2_mskcc_batch = assay(tcga_mskcc_batch[goi, c(idx1, idx2)])

map = data.frame(NAME = rownames(her2_gdc), DESCRIPTION = rep(NA, nrow(her2_gdc)))
her2_gdc = cbind(map, her2_gdc)
her2_xena = cbind(map, her2_xena)
her2_recount2 = cbind(map, her2_recount2)
her2_piccolo = cbind(map, her2_piccolo)
her2_mskcc_norm = cbind(map, her2_mskcc_norm)
her2_mskcc_batch = cbind(map, her2_mskcc_batch)

write.table(her2_gdc, file.path(gsea_dir, paste0("input_", tag, "_gdc_log2_tpm.txt")), 
            sep = "\t", quote=FALSE, row.names=FALSE, col.names=TRUE)
write.table(her2_xena, file.path(gsea_dir, paste0("input_", tag, "_XENA_log2_tpm.txt")), 
            sep = "\t", quote=FALSE, row.names=FALSE, col.names=TRUE)
write.table(her2_recount2, file.path(gsea_dir, paste0("input_", tag, "_Recount2_log2_tpm.txt")), 
            sep = "\t", quote=FALSE, row.names=FALSE, col.names=TRUE)
write.table(her2_piccolo, file.path(gsea_dir, paste0("input_", tag, "_piccolo_log2_tpm.txt")), 
            sep = "\t", quote=FALSE, row.names=FALSE, col.names=TRUE)
write.table(her2_mskcc_norm, file.path(gsea_dir, paste0("input_", tag, "_mskcc_norm_log2_tpm.txt")), 
            sep = "\t", quote=FALSE, row.names=FALSE, col.names=TRUE)
write.table(her2_mskcc_batch, file.path(gsea_dir, paste0("input_", tag, "_mskcc_batch_log2_tpm.txt")), 
            sep = "\t", quote=FALSE, row.names=FALSE, col.names=TRUE)


tag = "Lung"

goi= read.delim(file.path(s3_dir, "de_goi.txt"), header=FALSE, stringsAsFactors = FALSE )[,1]
luad_samples= read.delim(file.path(s3_dir,"luad_samples.txt"), header=FALSE, stringsAsFactors = FALSE )[,1]
lusc_samples= read.delim(file.path(s3_dir,"lusc_samples.txt"), header=FALSE, stringsAsFactors = FALSE )[,1]

idx1  = match( luad_samples, colnames(tcga_gdc))
idx2 = match(lusc_samples, colnames(tcga_gdc))
row_idx = match(goi, rownames(tcga_gdc))

table(is.na(idx1))
table(is.na(idx2))
table(is.na(row_idx))

lung_gdc = assay(tcga_gdc[goi, c(idx1, idx2)])
lung_xena = assay(tcga_xena[goi, c(idx1, idx2)])
lung_recount2 = assay(tcga_recount2[goi, c(idx1, idx2)])
lung_piccolo = assay(tcga_piccolo[ goi, c(idx1, idx2)])
lung_mskcc_norm = assay(tcga_mskcc_norm[goi, c(idx1,idx2)])
lung_mskcc_batch = assay(tcga_mskcc_batch[goi, c(idx1, idx2)])

map = data.frame(NAME = rownames(lung_gdc), DESCRIPTION = rep(NA, nrow(lung_gdc)))
lung_gdc = cbind(map, lung_gdc)
lung_xena = cbind(map, lung_xena)
lung_recount2 = cbind(map, lung_recount2)
lung_piccolo = cbind(map, lung_piccolo)
lung_mskcc_norm = cbind(map, lung_mskcc_norm)
lung_mskcc_batch = cbind(map, lung_mskcc_batch)

write.table(lung_gdc, file.path(gsea_dir, paste0("input_", tag, "_gdc_log2_tpm.txt")), 
            sep = "\t", quote=FALSE, row.names=FALSE, col.names=TRUE)
write.table(lung_xena, file.path(gsea_dir, paste0("input_", tag, "_XENA_log2_tpm.txt")), 
            sep = "\t", quote=FALSE, row.names=FALSE, col.names=TRUE)
write.table(lung_recount2, file.path(gsea_dir, paste0("input_", tag, "_Recount2_log2_tpm.txt")), 
            sep = "\t", quote=FALSE, row.names=FALSE, col.names=TRUE)
write.table(lung_piccolo, file.path(gsea_dir, paste0("input_", tag, "_piccolo_log2_tpm.txt")), 
            sep = "\t", quote=FALSE, row.names=FALSE, col.names=TRUE)
write.table(lung_mskcc_norm, file.path(gsea_dir, paste0("input_", tag, "_mskcc_norm_log2_tpm.txt")), 
            sep = "\t", quote=FALSE, row.names=FALSE, col.names=TRUE)
write.table(lung_mskcc_batch, file.path(gsea_dir, paste0("input_", tag, "_mskcc_batch_log2_tpm.txt")), 
            sep = "\t", quote=FALSE, row.names=FALSE, col.names=TRUE)



tag = "Kidney"

goi= read.delim(file.path(s3_dir, "de_goi.txt"), header=FALSE, stringsAsFactors = FALSE )[,1]
kirc_samples= read.delim(file.path(resdir,"kirc_samples.txt"), header=FALSE, stringsAsFactors = FALSE )[,1]
kirp_samples= read.delim(file.path(resdir,"kirp_samples.txt"), header=FALSE, stringsAsFactors = FALSE )[,1]

idx1  = match( kirp_samples, colnames(tcga_gdc))
idx2 = match(kirc_samples, colnames(tcga_gdc))
row_idx = match(goi, rownames(tcga_gdc))

table(is.na(idx1))
table(is.na(idx2))
table(is.na(row_idx))

kidney_gdc = assay(tcga_gdc[goi, c(idx1, idx2)])
kidney_xena = assay(tcga_xena[goi, c(idx1, idx2)])
kidney_recount2 = assay(tcga_recount2[goi, c(idx1, idx2)])
kidney_piccolo = assay(tcga_piccolo[ goi, c(idx1, idx2)])
kidney_mskcc_norm = assay(tcga_mskcc_norm[goi, c(idx1,idx2)])
kidney_mskcc_batch = assay(tcga_mskcc_batch[goi, c(idx1, idx2)])

map = data.frame(NAME = rownames(kidney_gdc), DESCRIPTION = rep(NA, nrow(kidney_gdc)))
kidney_gdc = cbind(map, kidney_gdc)
kidney_xena = cbind(map, kidney_xena)
kidney_recount2 = cbind(map, kidney_recount2)
kidney_piccolo = cbind(map, kidney_piccolo)
kidney_mskcc_norm = cbind(map, kidney_mskcc_norm)
kidney_mskcc_batch = cbind(map, kidney_mskcc_batch)

write.table(kidney_gdc, file.path(gsea_dir, paste0("input_", tag, "_gdc_log2_tpm.txt")), 
            sep = "\t", quote=FALSE, row.names=FALSE, col.names=TRUE)
write.table(kidney_xena, file.path(gsea_dir, paste0("input_", tag, "_XENA_log2_tpm.txt")), 
            sep = "\t", quote=FALSE, row.names=FALSE, col.names=TRUE)
write.table(kidney_recount2, file.path(gsea_dir, paste0("input_", tag, "_Recount2_log2_tpm.txt")), 
            sep = "\t", quote=FALSE, row.names=FALSE, col.names=TRUE)
write.table(kidney_piccolo, file.path(gsea_dir, paste0("input_", tag, "_piccolo_log2_tpm.txt")), 
            sep = "\t", quote=FALSE, row.names=FALSE, col.names=TRUE)
write.table(kidney_mskcc_norm, file.path(gsea_dir, paste0("input_", tag, "_mskcc_norm_log2_tpm.txt")), 
            sep = "\t", quote=FALSE, row.names=FALSE, col.names=TRUE)
write.table(kidney_mskcc_batch, file.path(gsea_dir, paste0("input_", tag, "_mskcc_batch_log2_tpm.txt")), 
            sep = "\t", quote=FALSE, row.names=FALSE, col.names=TRUE)

# create a cls files fir GSEA

kidney_cls = c(rep( "KIRP", length(kirp_samples) ), rep("KIRC", length(kirc_samples)))
write.table(t(kidney_cls), file.path(gsea_dir, "kidney_cls.cls"),
            sep = " ", quote=FALSE, row.names=FALSE, col.names=FALSE)


lung_cls = c(rep( "LUAD", length(luad_samples) ), rep("LUSC", length(lusc_samples)))
write.table(t(lung_cls), file.path(gsea_dir, "lung_cls.cls"),
            sep = " ", quote=FALSE, row.names=FALSE, col.names=FALSE)

breast_cls = c(rep( "Basal", length(basal_like_samples) ), rep("her2e", length(HER2_enriched_samples)))
write.table(t(breast_cls), file.path(gsea_dir, "Breast_cls.cls"),
            sep = " ", quote=FALSE, row.names=FALSE, col.names=FALSE)
```

### Supplemental Tables for GSEA analysis

```{r eval=FALSE}

rm(list=ls())
library(UpSetR)
gsea_dir = "~/HollandLabShared/Sonali/paper_with_Hamid_Siobhan/de_analysis/gsea_analysis"
setwd(gsea_dir)
fls = list.files(pattern = "_Basal_*")
fls = fls[ grep(".xls", fls)]
fls
breast_bigdf = lapply(fls, function(x) read.delim(x, header=T, stringsAsFactors=FALSE))
breast_pathways = lapply(breast_bigdf, function(x) x[which(x$FDR.q.val < 0.25), -c(2,3)] )
pipeline = c( rep("GDC", nrow(breast_pathways[[1]])), 
              rep("MSKCC Batch", nrow(breast_pathways[[2]])), 
              rep("MSKCC Norm", nrow(breast_pathways[[3]])), 
              rep("Piccolo", nrow(breast_pathways[[4]])), 
              rep("Recount2", nrow(breast_pathways[[5]])), 
              rep("Xena/Toil", nrow(breast_pathways[[6]]))
              )
breast_pathways = do.call(rbind, breast_pathways)
breast_pathways = cbind( pipeline , breast_pathways)


fls = list.files(pattern = "_KIRC_*")
fls = fls[ grep(".xls", fls)]
fls
kidney_bigdf = lapply(fls, function(x) read.delim(x, header=T, stringsAsFactors=FALSE))
kidney_pathways = lapply(kidney_bigdf, function(x) x[which(x$FDR.q.val < 0.25),  -c(2,3)] )
pipeline = c( rep("GDC", nrow(kidney_pathways[[1]])), 
              rep("MSKCC Batch", nrow(kidney_pathways[[2]])), 
              rep("MSKCC Norm", nrow(kidney_pathways[[3]])), 
              rep("Piccolo", nrow(kidney_pathways[[4]])), 
              rep("Recount2", nrow(kidney_pathways[[5]])), 
              rep("Xena/Toil", nrow(kidney_pathways[[6]]))
)
kidney_pathways = do.call(rbind, kidney_pathways)
kidney_pathways = cbind( pipeline , kidney_pathways)

fls = list.files(pattern = "_LUAD_*")
fls = fls[ grep(".xls", fls)]
fls
lung_bigdf = lapply(fls, function(x) read.delim(x, header=T, stringsAsFactors=FALSE))
lung_pathways = lapply(lung_bigdf, function(x) x[which(x$FDR.q.val < 0.25),  -c(2,3)] )
pipeline = c( rep("GDC", nrow(lung_pathways[[1]])), 
              rep("MSKCC Batch", nrow(lung_pathways[[2]])), 
              rep("MSKCC Norm", nrow(lung_pathways[[3]])), 
              rep("Piccolo", nrow(lung_pathways[[4]])), 
              rep("Recount2", nrow(lung_pathways[[5]])), 
              rep("Xena/Toil", nrow(lung_pathways[[6]]))
)
lung_pathways = do.call(rbind, lung_pathways)
lung_pathways = cbind( pipeline , lung_pathways)

table(lung_pathways[,1])
table(kidney_pathways[,1])
table(breast_pathways[,1])

write.table(lung_pathways, file.path(gsea_dir, "supp_table_gsea_lung.txt"), 
           sep = "\t", quote=FALSE, row.names=FALSE, col.names=TRUE)

write.table(kidney_pathways, file.path(gsea_dir, "supp_table_gsea_Kidney.txt"), 
           sep = "\t", quote=FALSE, row.names=FALSE, col.names=TRUE)


write.table(breast_pathways, file.path(gsea_dir, "supp_table_gsea_Breast.txt"), 
           sep = "\t", quote=FALSE, row.names=FALSE, col.names=TRUE)
```

## Barplots 

In the code below, we create side-by-side barplots to visualize DEGs and pathways found across all pipelines.
These are included as supplemental figures in our manuscripts.

```{r eval =FALSE}
library(ggplot2)
axis_text_size = 25
axis_title_size=25
legend_text_size=25


path_levs = c( "Pathway Found by all 5 pipelines", "Pathway Not Found by all 5 pipelines")
basal_df <- data.frame(dose=path_levs,
                       len=c( 17, 22), 
                       analysis=rep("Breast", 2))
luad_df <- data.frame(dose=path_levs,
                      len=c( 52, 42), 
                      analysis=rep("Lung", 2))
kirc_df <- data.frame(dose=path_levs,
                      len=c(193, 107), 
                      analysis=rep("Kidney", 2))
bar_df = rbind(basal_df, luad_df, kirc_df)
bar_df$dose = factor(bar_df$dose, levels = path_levs)


p5 =ggplot(data=bar_df, aes(x=analysis, y=len, fill=dose)) +
  geom_bar(stat="identity", position=position_dodge(), width = 0.8)+
  geom_text(aes(label=len), vjust=1.6, color="black",
            position = position_dodge(0.9), size=10)+
  scale_fill_brewer(palette="Dark2")+
  ylab("No of enriched pathways") + xlab("") +
  theme_bw(base_family="Helvetica") + 
  theme(legend.title = element_blank(), 
        legend.position="bottom", legend.direction="vertical", 
        legend.text = element_text(size = legend_text_size), 
        axis.text.x = element_text(colour = "black", size=axis_text_size), 
        axis.text.y = element_text(colour = "black", size=axis_text_size), 
        axis.title.x = element_blank(), 
        axis.title = element_text(size= axis_title_size)) 


# over-represented pathways
path_levs = c( "Pathways Found by all 5 pipelines", "Pathways Not Found by all 5 pipelines")
basal_df <- data.frame(dose=path_levs,
                       len=c( 8, 8), 
                       analysis=rep("Breast", 2))
luad_df <- data.frame(dose=path_levs,
                      len=c( 37, 10), 
                      analysis=rep("Lung", 2))
kirc_df <- data.frame(dose=path_levs,
                      len=c(58, 20), 
                      analysis=rep("Kidney", 2))
bar_df = rbind(basal_df, luad_df, kirc_df)
bar_df$dose = factor(bar_df$dose, levels = path_levs)


p6 =ggplot(data=bar_df, aes(x=analysis, y=len, fill=dose)) +
  geom_bar(stat="identity", position=position_dodge(), width = 0.8)+
  geom_text(aes(label=len), vjust=1.6, color="black",
            position = position_dodge(0.9), size=10)+
  scale_fill_brewer(palette="Dark2")+
  ylab("No of over-represented pathways") + xlab("") +
  theme_bw(base_family="Helvetica") + 
  theme(legend.title = element_blank(), 
        legend.position="bottom", legend.direction="vertical", 
        legend.text = element_text(size = legend_text_size), 
        axis.text.x = element_text(colour = "black", size=axis_text_size), 
        axis.text.y = element_text(colour = "black", size=axis_text_size), 
        axis.title.x = element_blank(), 
        axis.title = element_text(size= axis_title_size)) 



# differentially expressed genes. 
path_levs = c( "DEGs Found by all 5 pipelines", "DEGs Not Found by all 5 pipelines")

basal_df <- data.frame(dose=path_levs,
                       len=c( 802, 701), 
                       analysis=rep("Breast", 2))

luad_df <- data.frame(dose=path_levs,
                      len=c( 1806, 387), 
                      analysis=rep("Lung", 2))

kirc_df <- data.frame(dose=path_levs,
                      len=c(2020, 479), 
                      analysis=rep("Kidney", 2))

bar_df = rbind(basal_df, luad_df, kirc_df)
bar_df$dose = factor(bar_df$dose, levels = path_levs)

p7 =ggplot(data=bar_df, aes(x=analysis, y=len, fill=dose)) +
  geom_bar(stat="identity", position=position_dodge(), width = 0.8)+
  geom_text(aes(label=len), vjust=1.6, color="black",
            position = position_dodge(0.9), size=10)+
  scale_fill_brewer(palette="Dark2")+
  ylab("No of differentially expressed genes") + xlab("") +
  theme_bw(base_family="Helvetica") + 
  theme(legend.title = element_blank(), 
        legend.position="bottom", legend.direction="vertical", 
        legend.text = element_text(size = legend_text_size), 
        axis.text.x = element_text(colour = "black", size=axis_text_size), 
        axis.text.y = element_text(colour = "black", size=axis_text_size), 
        axis.title.x = element_blank(), 
        axis.title = element_text(size= axis_title_size)) 

pdf("side_by_side_barplots.pdf", width = 7, height = 7)
print(p5)
print(p6)
print(p7)
dev.off()
```

